---
title: Teams of complementary problem solvers outperform equally skilled individuals
output:
  bookdown::pdf_document2: {}
  bookdown::html_document2: {}
---
```{r inherited-diversity-config, include=FALSE}
library(knitr)
library(crotchet)
opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, fig.height=4, fig.width=4)
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "inherited-diversity.R"))
read_graphviz_chunk("team-structures", package = "gems", new_name = "gems-team-strategies")
```
```{r inherited-diversity}
```

# Introduction {-}

People can work together in teams to solve all sorts of problems. What is the
impact of teamwork on the problem solving abilities of the people in the team?
The purpose of this research was to test the conditions under which teamwork
might improve individual problem solving ability.

There are many ways to measure the effectiveness of teamwork. A common method
is to compare the performance of teams to individuals working alone. When
compared in this way, studies have often failed to find any evidence that
teamwork provides an improvement over individual problem solving.

One of the reasons teamwork is reflected rather poorly in the experimental
literature may be due to limits on experimental methods. In order to conduct an
experiment in which both teams and individuals are administered the same task,
many tasks are necessarily excluded because they can not be equally attempted
using both teams and individuals. To use a clear example: if the problem
solving task required moving heavy objects, teamwork has clear advantages, but
no one would be surprised by these results.

Teams can clearly solve many problems that are insurmountable for individuals.
But when both teams and individuals attempt tasks that individuals can complete
alone, teamwork is rarely found to be an improvement over individual
performance.

A notable exception to this trend was a study by @Bahrami:2010jl who found
that two heads were better than one in a perceptual judgment task. In the
task, participants searched for an oddball target and responded with a
binary judgment corresponding to which of two displays contained the
target. If the two participants in each team disagreed about which display
contained the target, they were allowed to make a joint decision. Because
the participants made both individual and team responses on each trial,
the researchers were able to compare the performance of the team to the
performance of best individual in each team. The results showed a clear
benefit to teamwork over what was would have been achieved by the
best individual working alone.

Teamwork has been shown to improve performance on search tasks.

A difficulty that arises when comparing experiments done on teamwork
is in how to properly control for time. The typical approach is to control
for **calendar hours**. For example, teams and individuals are both given
an hour to complete the same task. However, controlling for calendar hours does
not scale with team size. Individuals and teams are both given an hour,
regardless of the size of the team.

In this research we controlled for labor hours.







with team size. A single individual can be compare

Controlling for calendar hours
allows researchers to test questions



are compared relative to the same number of labor hours

for 


variability in how 

The reason teamwork


Part of the reason teamwork is reflected poorly in the experimental
literature is that there are many tasks that teams are good


The question of what makes for effective teamwork has many candidate

Seeking the recipe for successful teamwork has 

# Methods {-}

```{r gems-team-strategies, engine="dot", out.width="40%"}
```

# Version 1 {-}

```{r pilot}
```

```{r gabors-v1, fig.cap="Gabor space for version 1 of the experiment."}
draw_image("v1_SimpleHillGems", package = "gems")
```

```{r training-v1, fig.width=10, fig.height=8}
gridExtra::grid.arrange(
  read_image("v1_welcome_orientation", package = "gems"),
  read_image("v1_training_orientation", package = "gems"),
  read_image("v1_welcome_spatial_frequency", package = "gems"),
  read_image("v1_training_spatial_frequency", package = "gems"),
  nrow = 2
)
```

```{r pilot-scores, fig.width=6}
ggplot(Pilot) +
  aes(trial, score, color = instructions) +
  geom_line(aes(group = interaction(subj_id, run, landscape_ix)),
            size = 0.1) +
  geom_line(stat = "summary", fun.y = "mean", size = 1) +
  facet_wrap("training_label", scales = "free_x") +
  t_$theme +
  theme(legend.position = "bottom")
```

```{r pilot-training, fig.width=8, fig.height=8}
data("v1_OrientationBias")
data("v1_SpatialFrequencyBias")

orientation_bias <- wireframe(score ~ x * y, xlab = "ori", ylab = "sf", data = v1_OrientationBias)
spatial_frequency_bias <- wireframe(score ~ x * y, xlab = "ori", ylab = "sf", data = v1_SpatialFrequencyBias)

OrientationTraining <- Pilot %>%
  filter(instructions == "orientation", landscape_name == "OrientationBias")
SpatialFrequencyTraining <- Pilot %>%
  filter(instructions == "spatial_frequency", landscape_name == "SpatialFrequencyBias")

training_plot <- ggplot() +
  aes(x = current_x, y = current_y, group = subj_id) +
  geom_segment(aes(xend = selected_x, yend = selected_y, group = interaction(subj_id, run))) +
  annotate("point", x = 5, y = 5, shape = 4) +
  annotate("point", x = 50, y = 50, shape = 4) +
  t_$theme

gridExtra::grid.arrange(
  orientation_bias,
  (training_plot %+% OrientationTraining) +
    aes(color = subj_id) +
    ggtitle("Orientation Training"),
  spatial_frequency_bias,
  (training_plot %+% SpatialFrequencyTraining) +
    aes(color = subj_id) +
    ggtitle("Spatial Frequency Training"),
  nrow = 2
)
```

```{r simple-hill-landscape-v1}
data("v1_SimpleHill")
wireframe(score ~ x * y, data = v1_SimpleHill, xlab = "ori", ylab = "sf")
```

```{r pilot-test, fig.width=10}
PilotTest <- Pilot %>%
  filter(training == "test")

ggplot(PilotTest) +
  aes(current_x, current_y, color = instructions) +
  geom_segment(aes(xend = selected_x, yend = selected_y, group = subj_id)) +
  annotate("point", x = 5, y = 5, shape = 4) +
  annotate("point", x = 50, y = 50, shape = 4) +
  facet_wrap("landscape_label", nrow = 1) +
  t_$scale_x_orientation +
  t_$scale_y_spatial_frequency +
  t_$scale_color_instructions +
  t_$coord_landscape +
  t_$theme +
  theme(legend.position = "bottom")
```

```{r, eval=FALSE}
PilotSimpleHill <- filter(PilotTest, landscape_ix == 1)

custom_linetypes <- c(orientation='dot', spatial_frequency='solid')

plot_ly(PilotSimpleHill, x = ~selected_x, y = ~selected_y, z = ~score, color = ~subj_id, type = 'scatter3d',
        linetype = ~instructions, mode = 'lines+markers', split = "subj_id",
        opacity = 1, line = list(width = 2), marker = list(size=3), linetypes=custom_linetypes) %>%
    layout(
      scene = list(
        xaxis = list(title = "Orientation"),
        yaxis = list(title = "Spatial frequency"),
        zaxis = list(title = "Score"),
        camera = list(eye = list(x = -1, y = -1, z = 1))
      )
    )
```

# Version 2 {-}

```{r gems}
```

```{r gabors-v2, fig.cap="Gabor space for Version 2 of the experiment."}
draw_image("v2_SimpleHillGems", package = "gems")
```

```{r training-v2, fig.width=10, fig.height=8}
gridExtra::grid.arrange(
  read_image("v2_welcome_orientation", package = "gems"),
  read_image("v2_training_orientation", package = "gems"),
  read_image("v2_welcome_spatial_frequency", package = "gems"),
  read_image("v2_training_spatial_frequency", package = "gems"),
  nrow = 2
)
```

## Scores {-}

```{r gems-scores}
GemsScores <- Gems %>%
  group_by(instructions, training, landscape_ix, trial) %>%
  summarize(score = mean(score)) %>%
  ungroup()

ggplot(Gems) +
  aes(trial, score, color = instructions) +
  geom_line(aes(group = interaction(subj_id, landscape_ix)), size = 0.2) +
  geom_line(stat = "summary", fun.y = "mean", size = 2) +
  facet_wrap("training_label", scales = "free_x") +
  t_$theme +
  theme(legend.position = "bottom")
```

## Training {-}

```{r gems-training, fig.width=8, fig.height=8}
data("OrientationBias")
data("SpatialFrequencyBias")

orientation_bias <- wireframe(score ~ x * y, xlab = "ori", ylab = "sf", data = OrientationBias)
spatial_frequency_bias <- wireframe(score ~ x * y, xlab = "ori", ylab = "sf", data = SpatialFrequencyBias)

OrientationTraining <- Gems %>%
  filter(instructions == "orientation", landscape_name == "OrientationBias")
SpatialFrequencyTraining <- Gems %>%
  filter(instructions == "spatial_frequency", landscape_name == "SpatialFrequencyBias")

training_plot <- ggplot() +
  aes(x = current_x, y = current_y, xend = selected_x, yend = selected_y, group = subj_id) +
  geom_segment() +
  annotate("point", x = 5, y = 5, shape = 4) +
  annotate("point", x = 50, y = 50, shape = 4) +
  t_$theme

gridExtra::grid.arrange(
  orientation_bias,
  (training_plot %+% OrientationTraining) +
    aes(color = subj_id) +
    ggtitle("Orientation Training"),
  spatial_frequency_bias,
  (training_plot %+% SpatialFrequencyTraining) +
    aes(color = subj_id) +
    ggtitle("Spatial Frequency Training"),
  nrow = 2
)
```

## Test {-}

```{r gems-test, fig.width=8, fig.height=8}
data("SimpleHill")
simple_hill <- wireframe(score ~ x * y, data = SimpleHill, xlab = "Orientation", ylab = "Spatial frequency")

GemsTest <- filter(Gems, training == "test")
test_plot <- (training_plot %+% GemsTest) +
  aes(color = instructions) +
  facet_wrap("landscape_ix", nrow = 1) +
  theme(legend.position = "bottom")

gridExtra::grid.arrange(simple_hill, test_plot, nrow = 2)
```

## Walk {-}

```{r trials-walk, eval=FALSE}
library(gganimate)
library(geomnet)

Gems111 <- Gems %>%
  filter(subj_id == "GEMS111", landscape_ix == 1) %>%
  melt_trial_stims() %>%
  mutate(
    is_selected = (selected == gem_pos)
  ) %>%
  arrange(trial)

trials_plot <- ggplot(Gems111) +
  aes(gem_x, gem_y, frame = trial) +
  geom_segment(aes(current_x, current_y, xend = selected_x, yend = selected_y, group = 1, cumulative = TRUE)) +
  geom_circle(aes(x = current_x, y = current_y), radius = 0.35) +
  geom_point(aes(color = factor(is_selected), frame = trial))

gganimate(trials_plot)
```

# Discussion {-}
