---
title: Complementary inheritance in iterated problem solving
bibliography: references.bib
output:
  bookdown::pdf_document2:
    toc: false
    includes:
      in_header: templates/header.tex
abstract: |
  Humans have evolved a system of cultural inheritance that enables the
  accumulation of problem solving knowledge over generations. As a result,
  human problem solvers rarely have to start from scratch when solving new
  problems, and instead can benefit from solutions inherited from a previous
  generation. We investigated the impact of inheritance on problem solving,
  specifically the conditions under which inheritance might improve problem
  solving relative to solving the same problems individually. In Experiment 1,
  we simulated the effect of inheritance on problem solving using agent-based
  models, and demonstrate that teams of complementary problem solvers, who
  possess independent and complementary skills, are able to outperform equally
  skilled individuals. In Experiment 2, we tested these predictions in
  a behavioral experiment. The results illustrate one way in which our system
  of cultural inheritance may result in cumulative improvements in problem
  solving performance over generations.
---
```{r inherited-diversity-config, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results="hide",
               fig.height=4, fig.width=4)
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "inherited-diversity.R"))
read_chunk(file.path(chunks_dir, "peaks.R"))
```
```{r inherited-diversity}
```

# Introduction {-}

One of the fundamental processes in cultural evolution is the transmission of
problem solving knowledge from one generation to the next. Cultural
transmission is notoriously imperfect when compared to genetic transmission
[@Sperber:1996exc; @Claidiere:2010kd]. Yet the human ability to transmit
cultural information far exceeds the abilities of any non-human animals
[@Tomasello:1993vh; @Tennie:2009gaa]. Humans have evolved a number of abilities
that improve the fidelity of cultural transmission, including imitation,
teaching, and language [@Lewis:2012jg; @Dean:2012gv]. As a result of high
fidelity cultural transmission, human problem solvers rarely have to start from
scratch when solving new problems, and instead can benefit from discoveries
made by previous generations.

The purpose of this research was to investigate the consequences of inheriting
solutions from a previous generation on future problem solving. Specifically,
we were interested in the conditions under which problem solving organized
around inheritance may result in more effective problem solving than could be
achieved by solving the same problems individually. To test this, we compared
the performance of teams organized around inheritance, in which one person
hands off a partial solution to be continued by the next generation, to
individuals working alone.

Although the capacity to work together in teams to solve large problems is
evident in many scientific and technological domains, teamwork is not always
considered to be an effective means of improving individual problem solving
performance [@Propp:2003kj; @Tindale:1992ub]. When both teams and individuals
complete tasks that individuals can complete alone, teamwork does not usually
improve performance over what could have been accomplished individually.
Teamwork can introduce social factors such as free-riding
[@Holmstrom:1982cp], social loafing [@Latane:1979in], and bias
[@Lorenz:2011jn] that decrease motivation and group performance. Other
factors that reduce the effectiveness of teamwork include the increased cost of
communication among larger teams [@Brooks:1975mmm]. As a result, even if larger
teams are able to accomplish more than individuals, the contribution of each
individual team member may be smaller in a group setting [@Mao:2016ds].

One situation in which teamwork has been shown to reliably improve performance
over what could have been accomplished individually is in perceptual judgment tasks.
@Bahrami:2010jl found that teams of two individuals ("dyads") performed better
than individuals in a perceptual judgment task. On each trial, participants
searched for an oddball target and responded with a binary judgment. If the
dyad disagreed about which display contained the target, they were allowed to
confer and make a joint decision. Because participants made both individual and
team responses on the same trials, the researchers were able to compare the
performance of the team to the performance of best individual in each team. The
results showed a benefit to teamwork over what would have been achieved by the
best individual in each dyad completing the task alone [cf. @Sniezek:1989hn].
Follow-up studies have confirmed that dyads were able to improve team
performance through confidence matching. Confidence matching can occur through
settling on a common vocabulary to use when discussing uncertainty
[@Fusaroli:2012jf], but confidence matching also occurs when using
non-linguistic indicators of uncertainty [@Bang:2017cm].

In tasks involving confidence matching, teamwork improves performance
specifically when the members of the team have similar perceptual abilities.
When members of the team have different perceptual abilities, they are unable
to effectively combine their confidences in order to arrive at an optimal
decision [@Bahrami:2012dya; @DuncanReid:2017eo]. This finding contrasts with
other research that has argued for a benefit to diversity among team members
[@Jackson:2003hq; @Mathieu:2008gl]. For example, a common argument in economics
is that division of labor and specialization raises collective productivity
[@Becker:1992fd; @VonHippel:1994jc]. In problem solving contexts, diversity
among team members has been argued to be beneficial in tasks that require
searching a space of possible solutions, looking for the best ones. For these
types of problems, teams of equally skilled problem solvers may get stuck on
suboptimal solutions, where diverse teams are able to maintain a wider search
and in some cases can discover globally optimal solutions.

As evidence of this, @Hong:2004gt modeled groups of problem solvers, and showed
that groups of diverse problem solvers could outperform higher ability groups.
The reason was that the high ability groups were necessarily more similar in
problem solving strategy, and thus were more likely to become stuck on
suboptimal solutions to problems. The more diverse group was able to maintain
a wider search, and was less likely to get stuck.

We hypothesized that diversity among group members could become advantageous in
tasks involving the transmission of problem solutions from one generation to
the next. In these cases, the first generation of problem solvers in a group
can search for the best solution before passing it off to the next generation.
Because each generation has different skills, the first generation can start
the second generation off in a part of the solution landscape they would be
unlikely to reach on their own. As a result, there should be some classes of
problems for which team of diverse problem solvers may outperform than either
individual working alone.

In Experiment 1, we demonstrate the phenomena of complementary inheritance
using an agent-based model. In the model, problem solving agents traverse
a solution landscape looking for the best solutions. After a certain number
of guesses, they pass off their best guess to the next generation. This
next generation, who either has the same or different skills, continues the
work from the inherited location. We find that diverse teams of agents
outperform teams of equally skilled individuals, demonstrating complementary
inheritance.

In Experiment 2, we test these predictions in a behavioral experiment.
Participants completed a perceptual judgment task in which they searched for
the highest quality "space gems" (sinusoidal gabor patches) that varied in two
continuous dimensions. After a certain number of attempts, they passed their
best guess to the next generation. The next generation, who was trained to pay
attention to the same or different dimension of the gems, continues the search
from the inherited point. We predicted that teams trained on complementary
skills would outperform equally skilled individuals.

# Experiment 1: Agent-based model of complementary inheritance {-}

```{r peaks-setup}
```

In Experiment 1, we used an agent-based model to illustrate the phenomenon of
complementary inheritance. The model was designed to simulate the effect of
inheritance on problem solving performance. In the model, agents attempted to
solve a problem by searching a space of possible solutions looking for the best
ones. After a certain number of attempts, the best solutions discovered by each
agent were passed on to serve as the starting point for attempts by a new
agent. These second generation agents had either the same or different problem
solving skills. We explored whether skill diversity among generations of agents
could result in better problem solving performance than equally skilled individuals
given the same total number of attempts.

## Methods {-}

```{r peaks-methods}
```

Agents were assigned skill values corresponding to vision in two dimensions of
the solution landscape. All agents were allotted the same total amount of
skill, which was distributed between the two dimensions. Some agents had equal
skill in both dimensions, and others had unequal skill in either dimension
(Fig. \@ref(fig:ability-as-vision)A).

Agents with equal skill in both dimensions had a square range of vision, and
sampled solutions equally in all directions. Agents with unequal skills had
a rectangular range of vision, and were more likely to sample solutions along
one dimension over the other (Fig. \@ref(fig:ability-as-vision)B).

Agents with each type of skill distribution were arranged in pairs with other
agents, forming two-agent teams that varied in team skill distribution (Fig.
\@ref(fig:ability-as-vision)C). For identical teams, both agents had equal
skills along both dimensions. For diverse teams, each agent specialized in a
different skill.

Agents searched for better solutions using a simple hill-climbing procedure.
For each time step, agents randomly sampled a solution within their range of
vision, centered on their previous best solution. If the sampled solution was
better quality than their current best solution, i.e. it was higher on the
landscape, then the agent moved to the new position. For the next time step, a
solution was sampled relative to the new position. If the sampled solution
was not an improvement, agents did not move, and sampled from the previous best
solution again on the next time step.

(ref:ability-as-vision) Methods for agent-based models. **A.** Each agent was
assigned skill corresponding to vision in two dimensions. **B.** Skill values in
both dimensions were combined to form each agent's search areas, corresponding
to the range of solutions that could be sampled relative to any point on the
landscape. **C.** Teams of two agents differing in skill diversity. Identical
teams shared the same abilities in both dimensions. Complementary teams had
"equal but opposite" skills.

```{r ability-as-vision, fig.width=8, fig.height=8, fig.cap='(ref:ability-as-vision)'}
grid.arrange(
  gg_two_dimensions + ggtitle("A. Vision in two dimensions"),
  gg_search_rects + ggtitle("B. Search areas"),
  gg_differing_skills + ggtitle("C. Two person teams varying in vision distribution"),
  ncol = 1,
  heights = c(0.375, 0.25, 0.375)
)
```

## Results and Discussion {-}

```{r peaks-experiments}
```

The results of the simulations are shown in Fig. \@ref(fig:differing-skills).
When considering just the performance of the first generation agents, skill diversity was found
to negatively impact overall problem solving performance, such that agents with
unequal vision in both dimensions achieved lower performance scores than agents with equal skills,
`r exp1$gen1`. In contrast, after handing off to the second generation, skill
diversity significantly improved performance, `r exp1$gen2`, resulting in
a crossover effect, `r exp1$crossover`. That is, inheriting a lower quality solution
from an agent with unequal skills actually improved the performance of the next
generation, beyond what would have been achieved individually.

(ref:differing-skills) Results of agent-based model simulations.

```{r differing-skills, fig.width=8, fig.height=4, fig.cap="(ref:differing-skills)"}
limits <- seq(-100, 100, by = 10)
z <- expand.grid(x = limits, y = limits) %>%
     mutate(z = -x^2 - y^2)
simple_hill <- lattice::wireframe(z ~ x * y, data = z)

grid.arrange(simple_hill, gg_differing_skills_timeline, nrow = 1)
```

The reason skill diversity improved overall team performance even
though it decreased individual performance is shown in Fig.
\@ref(fig:differing-skills-walk). As skill diversity increases, first generation
agents proceed to parts of the solution landscape that individuals with equal
skill are unlikely to reach on their own. This means that second generation
agents, who inherit these positions, start off in a different part of the
solution space than they would have reached on their own. In this case, this
novel starting point positions second generation problem solvers so that they
are better able to reach the globally optimal solutions as compared to equally
skilled individuals.

(ref:differing-skills-walk) Grid positions of agent-based models by skill distribution.

```{r differing-skills-walk, fig.width=8, fig.height=2.5, fig.cap="(ref:differing-skills-walk)"}
gg_differing_skills_walk
```

The results of Experiment 1 demonstrate the phenomena of complementary
inheritance, by which inheritance over generations of diverse problem solvers
results in teams of problem solvers outperforming individuals trained on the
same skills and given the same number of attempts. In the simulations, this
improvement in performance is the result of the first generation putting later
generations in a part of the problem space that would be unlikely to be
achieved individually.

However, there are a number of limitations to this model. Most importantly,
problem solving is simulated as a random walk process, where solutions are
selected at random and only accepted if they improve performance. A more
ecologically valid test of whether inheritance can improve the performance of
diverse teams requires a task in which new solutions are selected strategically.
In Experiment 2, we test these predictions in a behavioral experiment.

# Experiment 2: Behavioral test of complementary inheritance {-}

In Experiment 2 we tested the predictions of the agent-based model presented in
Experiment 1 in a behavioral experiment. Participants were randomly assigned
to learn equal but independent skills in a perceptual discrimination task. In
the task, participants searched a space of possible solutions looking for the
best ones. We expected to find that teams of complementary problem solvers,
trained on independent skills, would be able to coordinate their labor so as to
improve problem solving performance beyond what either individual would have
achieved on their own.

## Methods {-}

```{r methods}
```

Participants completed a perceptual discrimination task where they were
presented displays of sinusoidal gabor patch stimuli that varied in orientation
and spatial frequency. We asked them to think of the gabors as precious gems on a
foreign planet. Their task was to learn which gems were more valuable than
others.

The value of each gem was determined by its orientation and spatial frequency.
The space of possible gems and their associated values can be visualized as
a three dimensional landscape. As participants selected gems across trials,
they effectively traversed the landscape for the highest peaks corresponding
to the most valuable gems. In order to select the most valuable gems on each
trial, participants had to learn which stimulus dimensions were the most
important.

On each trial, participants were presented six gems and they selected the one
they thought was the most valuable. The six gems presented on each trial were
sampled within a certain distance from the previously selected gem (Fig.
\@ref(fig:gems-trial)). The sampling procedure is described in the procedure,
below. As participants selected gems across trials, they effectively traversed
the solution landscape in a series of incremental steps.

(ref:gems-trial) Procedure. **A.** On each trial, participants were shown six gabor
patches referred to as "gems". Each gem was assigned a value based on its
orientation and spatial frequency. **B.** As participants selected gems across trials,
they effectively traversed the solution landscape, looking for the highest peaks.

```{r gems-trial, fig.width=6, fig.cap="(ref:gems-trial)"}
gridExtra::grid.arrange(
  crotchet::read_image("trial", package = "gems"),
  trial_plot,
  nrow = 1
)
```

Participants were trained to pay attention to either the orientation or the bar
width of the gems. The training procedure consisted of written instructions
and 30 training trials. The training procedure is detailed below.

Participants were also assigned to the first or second generation. First
generation participants started at the origin $(0_{ori},0_{bw})$ on each
landscape. Second generation participants were assigned to inherit from a first
generation participant. They inherited the location on the landscape
$(x_{ori},y_{bw})$ of the gem selected by the first generation participant after
a certain number of trials. **There was no interaction between first and second
generation participants other than that second generation participants started
each landscape off in a different part of the landscape than their first
generation predecessors.**

(ref:gems-team-structures) Conditions in Experiment 2. Participants were
trained to pay attention to either orientation (ori) or bar width (bw), and
assigned to generations. First generation participants (top row) completed the
task as **isolated** individuals. Second generation participants were assigned
to inherit from a first generation participant trained on the same dimension
(**congruent**, middle row), or on different dimensions (**complementary**,
bottom row). Left and right panels show counterbalanced conditions.

```{r gems-team-structures, fig.width=8, fig.cap="(ref:gems-team-structures)"}
crotchet::draw_graphviz("team-structures-counterbalanced", package = "gems")
```

In our analyses, we compared three strategies for finding the most valuable gems
(Fig. \@ref(fig:gems-team-structures)). Congruent teams were made of
participants trained to pay attention to the same stimulus dimension---either
orientation or bar width. Complementary teams were made of participants trained
on different stimulus dimensions. We also compared the performance of
individuals trained on a single dimension who completed the same total number of
trials on their own. We hypothesized that complementary teams of problem solvers
would be more efficient at traversing the solution landscapes than congruent
teams or isolated individuals trained to discriminate the same stimulus
dimension.

### Materials {-}

**Gabors.** We created a $71 \times 71$ grid of `r methods$n_gabors_in_landscape` gabor
patches that varied in orientation and spatial frequency. Orientations were sampled at regular
intervals from `r methods$min_ori`° to `r methods$max_ori`°. Spatial frequencies
were sampled at geometric intervals from `r methods$min_sf` cycles/pixel to
`r methods$max_sf` cycles/pixel. A geometric progression was used because high
spatial frequencies are harder to distinguish than low spatial frequencies.

**Landscapes.** The three different solution landscapes we used are show in Fig.
\@ref(fig:landscapes). There were two training landscapes, and one test
landscape. The training landscapes were constructed so that the value of the
gems was determined completely by one dimension (either orientation or spatial
frequency) and not the other. For the training landscapes, the
peak value was set at 50 units in either stimulus dimension.

Values in the test landscape were determined by both orientation and spatial
frequency, corresponding to a simple hill landscape with a single peak at `(50,50)`.
The simple hill landscape was symmetric with regard to training condition,
such that both instructions conditions were equally informative about the values
of the gems in the test landscape.

(ref:landscapes) Landscapes. Participants assigned to the
orientation or bar width condition completed the training trials on the
corresponding landscape. All participants completed the test trials on the same
simple hill landscape.

```{r landscapes, fig.width=8, fig.cap="(ref:landscapes)"}
gridExtra::grid.arrange(
  orientation_bias_landscape,
  spatial_frequency_bias_landscape,
  simple_hill_landscape,
  nrow = 1
)
```

**Survey.** After finishing the main experiment, participants completed a short survey. We
asked the participants to recall the instructions they received, to describe how
they approached the experiment, and to rate the extent to which they thought the
gems and scores were random. We were particularly interested in if the
participants felt they were moving around space of possible gems of their own
volition.

### Procedure {-}

Participants were randomly assigned to training condition (orientation, bar
width) and generation (first, second). The researcher told participants
they would be playing a game where they played as a space explorer who is trying
to find valuable gems on a foreign planet. Their goal was to learn to
distinguish the valuable gems from the cheap ones. Then participants read the
instructions corresponding to their assigned training condition, and completed a
block of 30 training trials. After the training trials, they completed four
blocks of 40 test trials. When finished, participants completed the
post-experiment survey. The experiment took around 30 minutes to complete.

**Training trial procedure.** Participants were randomly assigned to one of two
training conditions. In the orientation condition, participants read in their
instructions that the value of each gem was determined by the orientation of the
stripes. In the bar width condition, they read that the value was determined by
the width of the bars.

After reading the instructions, participants completed a block of 30 training
trials. Across the 30 trials, participants traversed a landscape corresponding
to their instructions condition. For example, a participant who was instructed
to pay attention to bar width completed the training trials on the landscape
that was biased so that the values of the gems were dependent completely on
spatial frequency and not on orientation.

On the first training trial, participants were shown six gems sampled from
around the origin $(0_{ori},0_{bw})$ of the stimulus space. On subsequent
training trials, gems were sampled from around the position of the gem
$(x_{ori},y_{bw})$ selected on the previous trial.

In the training trials, after selecting a gem, participants were shown the
values of all six gems, and had to click on the most valuable gem in the
display to continue. This second response was required to encourage
participants to pay attention to which gabors were the most valuable so they
could more easily learn to select the most valuable gem on each trial. If the
participant had correctly selected the most valuable gem in their first
response, they were congratulated for making the correct choice, and told to
click on the gem they selected again.

**Test trial procedure.** After completing the training trials, participants
completed four blocks of 40 test trials. All test trials were conducted on the
same simple hill landscape, defined above. Participants traversed the simple
hill landscape four times, once for each block of 40 trials. First generation
participants started each block with 6 gems sampled from around the origin
$(0_{ori},0_{bw})$. Second generation participants participants started off at
the location of a first generation participant after 20 trials,
$(x_{ori},y_{bw})$. First and second generation participants were yoked over
blocks such that second generation participants inherited from the same first
generation participant four times total, one for each block. Participants were
given a break after each block, and told that for the next block, they may
encounter gems that they had not seen before.

In the test trials, after selecting a gem, participants were shown only the
value of the gem they selected, and not the values of the gems they did not
select. The value were transformed to be relative to the value of the
gem selected on the previous trial such that positive numbers indicated
an increase in overall score.

### Participants {-}

Participants (N=`r methods$n_participants`) were undergraduates at the University of Wisconsin-Madison who
completed the experiment in exchange for course credit. We collected
data for `r methods$n_first_gen` first generation participants, and `r methods$n_second_gen` second generation
participants. There were twice as many second generation participants because
each first generation participant was yoked to two second generation
participants, one in each training condition (see Fig. \@ref(fig:gems-team-structures)).

## Results {-}

```{r results}
```

In our analyses, we report performance on each trial and across trials. For
performance on each trial, we analyzed which of the six gems presented on each
trial was selected based on the attributes of that gem (orientation and bar
width). For performance across trials, we analyzed how participants traversed
the solution landscape, in particular their distance to the optimal values in
each dimension.

Before comparing the performance of complementary, congruent, and isolated strategies,
we first verified the success of the training procedure. After showing the that
the training procedure was successful, we turn to analyzing performance in the
test trials.

### Training {-}

The purpose of the training procedure was to induce participants to learn to pay
attention to independent dimensions of the gabor stimuli corresponding
to their training condition. To verify the success of the training procedure, we
measured the extent to which participants selected stimuli in each training
trial corresponding to their training condition.

Participants in both training conditions traversed the same stimulus space
in orthogonal directions (Fig. \@ref(fig:gems-training)A). As expected,
participants selected gems on each trial that were closer to
the most valuable gems on their assigned training landscape. The relationship was
linear, `r r_$training_distance_linear_trial`, with most participants reaching the peak
of the training landscape within 30 trials.

Although participants in both training conditions proceeded in orthogonal
directions, they did not do so equally. Participants trained on bar width were
able to hone in on the optimal stimuli in fewer trials than participants
trained on orientation, `r r_$training_distance_linear_trial_v_instructions`
(Fig. \@ref(fig:gems-training)B).

(ref:gems-training) Results of training procedure. Participants were
randomly assigned to either the orientation or the bar width condition.
**A.** Grid positions. Participants in both training conditions started at
$(0_{ori},0_{bw})$ and traversed the same stimulus space in orthogonal
directions, each progressing toward the optimal stimulus value in the trained
dimension. **B.** Distance. Grid positions were converted to 1D distances
relative to the optimal value in each dimension. **C.** Score. Cumulative
increases in score over trials, indicating that participants were selecting
incrementally more valuable gems over trials.

```{r gems-training, fig.width=10, fig.height=4, fig.cap="(ref:gems-training)"}
gridExtra::grid.arrange(
  training_positions_plot + ggtitle("A. Grid position"),
  training_distance_plot + ggtitle("B. Distance"),
  training_scores_plot + ggtitle("C. Score"),
  nrow = 1
)
```

We also measured the perceptual sensitivity of participants to the different
stimulus dimensions on a trial-by-trial basis. For this analysis, we transformed
the stimulus dimensions of the gems displayed on each trial to be relative to
the optimal value along that dimension. For example, for a participant assigned
to the orientation training condition, gems with orientations at grid points
40 and 60 were both 10 units away from the optimal orientation at 50 units. These
relative values were then centered for each trial such that positive values
indicated that the gem was an improvement in score relative to the previous
position. Using these transformed values, we fit two models, one
predicting sensitivity to the orientation dimension, and another predicting
sensitivity to the bar width dimension. We included the instruction condition
of the participants in the models. We expected participants in either training
condition to be more sensitive to to their trained dimension than the untrained
dimension.

As expected, participants trained on either stimulus dimension selected
gems on each trial corresponding to their trained dimension and were
uninfluenced by variability in the untrained dimension, orientation:
`r r_$training_sensitivity_ori`, bar width: `r r_$training_sensitivity_sf`.

To compare the effectiveness of either training
condition, we transformed the sensitivities to be relative to training
condition. Unlike in the analysis of performance across trials reported above, our
trial-by-trial analysis did not reveal a difference between training conditions,
`r r_$training_sensitivity_trained_dimension_by_instructions`.

(ref:gems-training-sensitivities) Sensitivity to stimulus dimensions in the training
procedure. Each thin line is the average sensitivity for a single participant.
Thick lines correspond to the predictions of the hierarchical linear model.
**A.** Sensitivity to orientation and bar width separated by training condition.
**B.** Trained dimension collapsed over training condition.

```{r gems-training-sensitivities, fig.width=8, fig.height=3, fig.cap="(ref:gems-training-sensitivities)"}
gridExtra::grid.arrange(
  training_sensitivity_plot + ggtitle("A. Sensitivity to orientation and bar width"),
  training_sensitivity_trained_dimensions_plot +
    ggtitle("B. Trained dimension") +
    # removing y-axis title and legend because of compound plot
    theme(
      axis.title.y = element_blank(),
      legend.position = "none"
    ),
  nrow = 1,
  widths = c(0.66, 0.34)
)
```

The results of the training procedure indicate that participants were
successfully induced to discriminate independent dimensions of the gabor patch
stimuli. We now turn to our primary hypothesis of whether participants trained
on independent dimensions were able to coordinate their problem solving
behavior more effectively than participants trained on the same dimension.

### Performance {-}

In our first analysis, we ask whether individuals completing this problem
solving task benefited from inheritance, or whether they would have been better
off completing the same number of trials as isolated individuals. To test this,
we compared first generation performance after 40 trials in each of the four
test blocks to second generation performance after 20 trials. Since second
generation participants start off where a particular first generation
participant was after 20 trials, our analysis effectively asks whether first
generation individuals, after completing 20 trials, are better off passing
their partial solution on to the next generation, or if they are better off
completing the additional 20 trials themselves.

We focus on two separate but related scoring metrics: 2D distance,
corresponding to the Euclidean distance between the participants last selected
gem and the peak of the test landscape at $(50,50)$, and score, corresponding
to the height or value of the gem last selected. Each scoring metric can be
taken in absolute units, or in relative units. Relative distance and relative
score correspond to the difference between the starting point and the current
point in terms of distance and score, such that positive numbers mean the
participant was able to move closer to the target and to improve their overall
score.

After 20 trials, complementary problem solvers were able
to achieve slightly higher absolute scores than both isolated and congruent strategies,
although this difference was not statistically significant,
`r r_$strategies_scores_helmert_mod_main`. There was no residual difference
between isolated and congruent strategies, `r r_$strategies_scores_helmert_mod_resid`,
(Fig. \@ref(fig:strategies)A). Complementary problem solvers were
able to achieve higher relative scores than both isolated and congruent strategies,
although this difference was only marginal, `r r_$strategies_relative_scores_helmert_mod_main`.
Again, there was no residual difference between the isolated and congruent strategies,
`r r_$strategies_relative_scores_helmert_mod_resid` (Fig. \@ref(fig:strategies)B).

(ref:strategies) Hill climbing performance. **A.** Scores correspond to heights
on the landscape, and **C.** Distances correspond to Euclidean distances between
current location and peak in the landscape. **B.** and **D.** show these same
measures controlling for starting position.

```{r strategies, fig.width=7, fig.height=7, fig.cap="(ref:strategies)"}
gridExtra::grid.arrange(
  strategies_scores_plot + ggtitle("A. Final score"),
  strategies_relative_scores_plot + ggtitle("B. Relative score"),
  strategies_distance_plot + ggtitle("C. Distance to peak"),
  strategies_relative_distance_plot + ggtitle("D. Relative distance"),
  nrow = 2
)
```

### Inherited distance {-}

(ref:achieved-distance-to-1d) Performance relative to inherited position. Inherited distance
to trained peak is the distance toward the trained peak that was inherited, where larger numbers
indicate that the ancestor proceeded further to the 1D peak. Achieved distance shows the distance
moved toward the peak.

```{r achieved-distance-to-1d, fig.cap="(ref:achieved-distance-to-1d)"}
achieved_to_inherited_1d_plot
```

### Sensitivity {-}

In addition to measuring the impact of inheritance on final team performance,
we also analyzed whether complementary inheritance resulted in second generation
participants becoming more sensitive to either the trained or untrained dimensions.
We hypothesized that complementary problem solvers who start off in a part of the
solution landscape that would be unlikely to reached alone become exposed to
a wider variety of different gem stimuli than either a first generation individual
or an individual who inherits from a congruent ancestor.

To test this, we used the same sensitivity measure as reported in the Training
procedure above. The sensitivity measure reflects the extent to which participants
select gems in each trial corresponding to it's value along a particular dimension.
We expected all participants in the test trials to be more sensitive to their
trained dimension than their untrained dimension. Note that this is true even
though in the test landscape, both dimensions are weighted equally.

We also compared the sensitivity to trained and untrained dimensions across team
strategies. We hypothesized that complementary problem solvers, by being exposed
to a wider range of gems, might become more sensitive not only to their trained
dimension, but also their untrained dimension. To test this, we compared sensitivity
in trained and untrained dimensions across team conditions.

We found that a complementary strategy resulted in second generation participants
becoming more sensitive to their trained dimension, `r r_$trained_sensitivity_main`,
with no residual difference between isolated and congruent strategies, `r r_$trained_sensitivity_resid`.

We did not find any evidence that complementary inheritance affected overall
sensitivity, `r r_$overall_sensitivity_main` or sensitivity in the untrained
dimension, `r r_$untrained_sensitivity_main`.

(ref:sensitivities) Sensitivity to trained and untrained dimensions by team strategy.

```{r sensitivities, fig.width=8, fig.cap="(ref:sensitivities)"}
sensitivities_plot
```

## Discussion {-}

Experiment 2 was designed to test whether teams of participants trained on
different but complementary skills could work together to achieve performance
that was greater than either individual could have achieved on their own.
We found some support for our hypothesis that problem solving with inheritance can
result in an improvement in performance. In the experiment, second generation
problem solvers who inherited from a previous generation were able to
progress closer to the optimal peak than problem solvers working in isolation
or who inherit from a congruent problem solver trained on the same skills.
These results validate the main conclusions drawn in Experiment 1 in a behavioral
experiment.

# Discussion {-}

The ability to pass problem solving knowledge on to future generations is
fundamental to human problem solving ability [@Boyd:2011gp]. Yet the
inheritance of problem solving knowledge can be costly, and it may take years
to develop expertise [@Stout:2002iu]. Here we investigated whether inheritance
can in some cases result in team performance greater than what could be
achieved by individuals working alone. We hypothesized that complementary
inheritance, whereby individuals with independent but complementary skills
collaborate over generations, may result in an overall advantage to inheritance
in a problem solving context.

In Experiment 1 we used an agent-based model to illustrate the conditions under
which inheritance of complementary problem solving knowledge might prove
effective over alternative strategies. We found that teams of complementary
problem solvers can start future generations off in a part of the problem
space that they would be unlikely to reach on their own. In some cases, this
novel starting position puts problem solvers in a location where they are
better situated to find the globally optimal solution.

Although the results of the agent-based models were suggestive, they needed to
be verified using a more ecologically valid experiment. In Experiment 2, we
test the main predictions drawn from Experiment 1 about the conditions under
which complementary inheritance might prove effective. We found initial support
for our primary hypothesis that there are some problems for which teams of
complementary problem solvers can enable problem solving behaviors that no
individual could achieve on their own.

These results add to a growing body of literature that aims to understand why
and how human problem solvers are exceptionally skilled at inheriting and
adapting the tools and other solutions to problems that were originally
discovered by someone else. Here we demonstrate that inheritance over
generations may result in an improvement to problem solving performance, beyond
what could be achieved individually.
