---
title: Inherited knowledge
output:
  bookdown::pdf_document2:
    toc: no
  github_document:
    toc: yes
    toc_depth: 3
---
```{r config, include=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE)
read_chunk("R/inherited-instructions.R")

library(patchwork)
library(pander)
panderOptions("table.alignment.default", "left")

library(lme4)
library(AICcmodavg)
```
```{r inherited-instructions, include=FALSE}
```

# Introduction

## Research question

Can individuals pass on knowledge via language that enables future generations
to exceed previous generations?

# Methods

Participants had to learn to categorize visual stimuli on the basis of two
visual dimensions. The stimuli were sinusoidal gabor patches that varied in
orientation and spatial frequency. Participants were told to think of each gabor
patch as a "gem" and that their task was to find the most valuable gems. Finding
the most valuable gems required integrating along both the orientation and
spatial frequency dimensions.

Participants were assigned to Generation 1 or Generation 2. Generation 1
participants completed the task naively, without any instructions about what
makes some gems more valuable. After 1 block of trials, these participants wrote
down a description of what makes some gems more valuable so as to be
helpful to the next generation. Generation 2 participants inherit these
instructions, and read them before completing the trials. After 1 block
of trials, Generation 2 participants write down their description of
what makes some gems more valuable, and complete the second block of trials.

## Procedure

On each trial, participants are presented 6 gabor patches ("gems") that vary in
orientation and spatial frequency. They have to select one they think will be
most valuable. After selecting a gem, they receive feedback about its value
relative to the value of the gem they selected on the previous trial. For
example, a participant might learn that the gem they selected that was worth +3
more than their previous one. They are also shown their new total score, a number
between 0 and 100.

\newpage

# Results

## Scores

(ref:scores) Score over trials and blocks. Gray lines show simulated performance
of optimal and random walk agents.

```{r scores, fig.width=8, fig.height=4, fig.cap="(ref:scores)"}
scores_plot + final_scores_plot
```

## Reaction times

```{r rts, fig.width=5, fig.height=4}
rts_plot
```

## Instructions

```{r instructions-coded, fig.width=8, fig.height=3}
instructions_coded_plot + instructions_summarized_plot + plot_layout(nrow = 1)
```

```{r instructions-outcomes, fig.width=6, fig.height=4, fig.caption="Correlation between score in block 1 and instruction quality. Participants who generated better instructions tended to do better in the task."}
instruction_quality_and_performance_plot + instruction_quality_and_final_performance_plot
```

\newpage

## The effects of inherited instructions

### Are instructions better than no instructions?

Participants in Generation 2 will benefit from accurate instructions
in block 1 relative to naive participants in Generation 1, block 1.

```{r instructions-v-no-instructions}
instructions_versus_no_instructions_plot
```

\newpage

### Are instructions better than 1 block of training?

Participants in Generation 2 will benefit from accurate instructions
relative to how participants in Generation 1 perform _after_ generating
the instructions.

```{r instructions-v-one-block}
instructions_versus_one_block_plot
```
