---
title: Inherited knowledge
output:
  bookdown::pdf_document2:
    toc: no
---
```{r config, include=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
read_chunk("R/inherited-instructions.R")

library(patchwork)
library(pander)
```
```{r inherited-instructions, include=FALSE}
```

# Introduction

## Research question

Can individuals pass on knowledge via language that enables future generations
to exceed previous generations?

# Methods

## Procedure

On each trial, participants see 6 gabor patches ("gems") that vary in
orientation and bar width and they select the one they think is most valuable.

Unbeknownst to the participants, the 6 gabors they see on each trial
are sampled from within a certain radius in stimulus space of their last
trial.

Participants are assigned to Generation 1 or Generation 2.

- Generation 1 participants complete 2 blocks, then they type out instructions
  for the next generation, then they complete 2 more blocks.
- Generation 2 participants read the instructions from Generation 1, then
  complete 4 blocks.

# Results

## Generation 1

```{r gen1, include=FALSE}
```

```{r gen1-results, fig.width=8, fig.height=8}
gen1_positions_plot + gen1_scores_plot + gen1_distance_plot + plot_layout(nrow = 3)
```

```{r gen1-final, fig.width=6}
gen1_final_scores_plot + gen1_final_distances_plot
```

### Instructions

```{r instructions}
data("Instructions")
data("Survey")

panderOptions("table.alignment.default", "left")
left_join(Instructions, Survey[,c("subj_id", "instructions_2")]) %>%
  pander(split.cell = 25)
```

### Survey

```{r survey}
data("Survey")

predictable <- ggplot(Survey) +
  aes(predictable) +
  geom_histogram(binwidth = 1, color = "white") +
  labs(title = "How predictable or random did you\nfeel the values of the gems were?") +
  scale_x_continuous(breaks = 1:5) +
  coord_cartesian(xlim = c(0.5,5.5))

effort <- ggplot(Survey) +
  aes(effort) +
  geom_histogram(binwidth = 1, color = "white") +
  labs(title = "Please rate your overall effort\nin this experiment.") +
  scale_x_continuous(breaks = 1:5) +
  coord_cartesian(xlim = c(0.5,5.5))

predictable + effort
```

## Generation 2

1. How does the performance of Generation 1 participants in blocks 1 and 2
   compare to the performance of Generation 2 participants in blocks 1 and 2?
2. How does the performance of Generation 1 participants in blocks 3 and 4
   compare to the performance of Generation 2 participants in blocks 1 and 2?
