---
title: Complementary problem solving in diachronic collaboration
bibliography: references.bib
output:
  bookdown::pdf_document2:
    toc: false
  bookdown::html_document2: {}
---
```{r inherited-diversity-config, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE, results="hide",
               fig.height=4, fig.width=4)
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "inherited-diversity.R"))

library(crotchet)
read_graphviz_chunk("team-structures-instructions", package = "gems",
                    new_name = "gems-team-structures-instructions")
read_graphviz_chunk("team-structures-generations", package = "gems",
                    new_name = "gems-team-structures-generations")
```
```{r inherited-diversity}
```

# Introduction {-}

One of the fundamental processes in cultural evolution is the transmission of
problem solving knowledge from one generation to the next. Cultural transmission
is notoriously imperfect compared to, for example, genetic transmission. Yet the human
ability to transmit cultural information far exceeds the abilities of any
non-human animals. Humans have evolved a number of abilities that improve the
success of cultural transmission, including teaching, language, and other social
learning abilities. The ability to reproduce tools and other solutions to
problems over generations is particularly beneficial in that future generations
of problem solvers rarely have to start from scratch when solving new problems.
Instead, they start off with solutions that were inherited from a previous
generation. How does the starting point in a problem solving task affect future
problem solving?

Some starting points are more beneficial than others. A laptop is clearly more
beneficial than a calculator. But does the starting point affect the way in
which the problem is solved? If the starting point can improve future problem
solving performance, then our system of cultural inheritance may actually
contribute to a general improvement in human problem solving ability.

To answer this question, we had participants complete a problem solving task
either as first or second generation problem solvers. First generation problem
solvers attempted a problem solving task that simulated a hill climbing
procedure. All first generation problem solvers started off from the same
Second generation problem solvers attempted the same problem solving
task but starting at where the first generation problem solvers had been
at the halfway point.

A consequence of our system of cultural inheritance is that problem solvers
rarely have to start from scratch when solving new problems. Instead, they start
off with the solutions that were inherited from a previous generation.

There are many ways to measure the effectiveness of teamwork. A common method
is to compare the performance of teams to individuals working alone. When
compared in this way, studies have often failed to find any evidence that
teamwork provides an improvement over individual problem solving.

One of the reasons teamwork is reflected rather poorly in the experimental
literature may be due to limits on experimental methods. In order to conduct an
experiment in which both teams and individuals are administered the same task,
many tasks are necessarily excluded because they can not be equally attempted
using both teams and individuals. To use a clear example: if the problem
solving task required moving heavy objects, teamwork has clear advantages, but
no one would be surprised by these results.

Teams can clearly solve many problems that are insurmountable for individuals.
But when both teams and individuals attempt tasks that individuals can complete
alone, teamwork is rarely found to be an improvement over individual
performance.

A notable exception to this trend was a study by @Bahrami:2010jl who found
that two heads were better than one in a perceptual judgment task. In the
task, participants searched for an oddball target and responded with a
binary judgment corresponding to which of two displays contained the
target. If the two participants in each team disagreed about which display
contained the target, they were allowed to make a joint decision. Because
the participants made both individual and team responses on each trial,
the researchers were able to compare the performance of the team to the
performance of best individual in each team. The results showed a clear
benefit to teamwork over what was would have been achieved by the
best individual working alone.

Teamwork has been shown to improve performance on search tasks.

A difficulty that arises when comparing experiments done on teamwork
is in how to properly control for time. The typical approach is to control
for **calendar hours**. For example, teams and individuals are both given
an hour to complete the same task. However, controlling for calendar hours does
not scale with team size. Individuals and teams are both given an hour,
regardless of the size of the team.

In the experiment, we compare three strategies for solving the same problem.

# Methods {-}
```{r methods}
```

Participants completed a perceptual discrimination task where they were
presented displays of sinusoidal gabor patch stimuli that varied in orientation and
spatial frequency. We asked them to think of the gabors as "gems"
from a foreign planet. Their task was to learn which gems were
more valuable than others.

The value of each gem was determined by its orientation and spatial frequency.
The space of possible gems and their associated values can be visualized as a
three dimensional landscape. As participants selected gems across trials, they
effectively explored different parts of this landscape, looking for the highest
peaks, corresponding to the most valuable gems. In order to detect
the most valuable gems on each trial, participants had to learn which stimulus
dimensions were the most important.

On each trial, participants selected one of six gabor patches they thought
was the most valuable. The six gabors presented on each trial were sampled
within a certain distance of previously selected gem. In order to find the most
valuable gems, participants had to navigate the solution landscape in a series
of incremental steps.

Participants were trained to pay attention to either the orientation or the bar
width of the gems. The training procedure consisted of written instructions
and 30 training trials. The training procedure is detailed below.

Participants were also assigned to the first or second generation. First
generation participants started at the origin `(0,0)` on each landscape. Second
generation participants were randomly assigned to inherit from a first
generation participant. They inherited the location on the landscape of the gem
selected by the first generation participant. There was no interaction between
first and second generation participants other than that second generation
participants started each landscape off in a different part of the landscape
than their predecessors.

In our analyses, we compare three problem solving strategies. Congruent teams
were made of participants trained to pay attention to the same stimulus
dimension---either orientation or bar width. Complementary teams were made of
participants trained on different stimulus dimensions. As a control, we also
compared the performance of individuals trained on a single dimension who
completed the same total number of trials on their own.

(ref:gems-team-structures-instructions) Team structures by instructions.
Participants were trained to pay attention to either orientation or bar width,
and assigned to generations. First generation participants completed the task as
isolated individuals. Second generation participants were assigned to inherit
from a first generation participant.

```{r gems-team-structures, fig.width=8, fig.cap="(ref:gems-team-structures-instructions)"}
gridExtra::grid.arrange(
  crotchet::read_graphviz("team-structures-instructions", package = "gems"),
  crotchet::read_graphviz("team-structures-generations", package = "gems"),
  nrow = 1, widths = c(0.45, 0.55)
)
```

## Materials {-}

### Gabors {-}

We created a 71x71 grid of `r methods$n_gabors_in_landscape` gabor patches that
varied in orientation and spatial frequency. A sample of the space is shown in
Fig. \@ref(fig:gabors). Orientations were sampled at regular intervals from
`r methods$min_ori`° to `r methods$max_ori`°. Spatial frequencies were sampled at
geometric intervals from `r methods$min_sf` cycles/pixel to `r methods$max_sf`
cycles/pixel. A geometric progression was used because high spatial frequencies
are harder to distinguish than low spatial frequencies.

(ref:gabors) Sampled space of possible gabor patches. Gabor patches varied
in orientation and spatial frequency.

```{r gabors, fig.cap="(ref:gabors)"}
draw_image("SimpleHillGems", package = "gems")
```

### Solution landscapes {-}

We created three different solution landscapes, shown in Fig.
\@ref(fig:landscapes). There were two different training
landscapes, and one test landscape. The training landscapes were constructed so
that the value of a gem was determined completely by one dimension (either
orientation or spatial frequency) and not influenced by the other dimension. For
these landscapes, the peak value was set at 50 units in either stimulus dimension.
Values in the test landscape were determined by both orientation and spatial
frequency, corresponding to a simple hill landscape with a single peak at `(50,50)`.

(ref:landscapes) Training and test landscapes. Participants assigned to the
orientation or bar width condition completed the training trials on the
corresponding landscape. All participants completed the test trials on the same
simple hill landscape.

```{r landscapes, fig.width=8, fig.cap="(ref:landscapes)"}
gridExtra::grid.arrange(
  orientation_bias_landscape,
  spatial_frequency_bias_landscape,
  simple_hill_landscape,
  nrow = 1
)
```

### Post experiment survey {-}

After finishing the main experiment, participants completed a short survey.
We asked the participants about the instructions they received, how they
approached the experiment, and to what extent they thought the gems and
scores were random. We were particularly interested in if the
participants felt they were moving around space of possible gabors
of their own volition. The questions are listed in Table \@ref(tab:survey-questions).

(ref:survey-questions) Questions asked in the post-experiment survey.

```{r survey-questions, results='asis'}
data("Survey")
data_frame(question_text = colnames(Survey)) %>%
  mutate(question_id = c("timestamp", "subj_id", "computer", "age", "gender", paste0("Q", 1:9))) %>%
  filter(grepl("Q", question_id)) %>%
  select(`#` = question_id, `Text` = question_text) %>%
  kable(caption="(ref:survey-questions)") %>%
  column_spec(2, width = "45em")
```

## Procedure {-}

Participants read the instructions corresponding to their randomly assigned
training condition, and completed a block of 30 training trials. After the
training trials, they completed four blocks of 40 test trials. After all four
blocks of test trials, participants completed the post-experiment survey. The
experiment took around 30 minutes to complete.

### Training {-}

Participants were randomly assigned to one of two training conditions. In the
orientation condition, participants were told that the value of each gem
was determined by the orientation of the stripes. In the bar width condition,
they were told the value was determined by the width of the bars. The exact
instructions are shown in Fig. \@ref(fig:training).

After reading the instructions, participants completed a block of 30 training
trials on a landscape corresponding to their instructions condition. For
example, a participant who was instructed to pay attention to bar width
completed the training trials on the landscape that was biased so that the
values of the gems were dependent completely on spatial frequency and not on
orientation.

For the training trials, after each choice, participants were shown the values
of all the gems in the display, and had to click on the most valuable
gem to continue to the next trial. This second response was to encourage
participants to learn which gabors were the most valuable. If the participant
had correctly selected the most valuable gem in their first response, they were told to
click on the gem they selected again. An example training trial is shown
in Fig. \@ref(fig:training-trial).

### Test {-}

After completing the training trials, participants completed four blocks of 40
trials. All test trials were conducted on the test landscape, described above.
Participants traversed the same landscape four times.

First generation participants started off at the origin `(0,0)` for each block. Second
generation participants started off at the location of a first
generation participant after 20 trials. First and second generation participants
were yoked over blocks such that second generation participants inherited from
the same first generation participant a total of four times, one for each
block.

## Participants {-}

Participants were undergraduates at the University of Wisconsin-Madison who
completed the experiment in exchange for course credit. We decided to collect
data for 40 first generation participants, and 80 second generation
participants. There were twice as many second generation participants because
each first generation participant was yoked to two second generation
participants, one in each training condition.

# Results {-}
```{r results}
```

## Training {-}

First and second generation participants completed the same training procedure.
In the training procedure, participants traversed either the oriention bias landscape
or the spatial frequency bias landscape. As expected, participants in either training
condition moved through the same stimulus space in orthogonal directions
(Fig. \@ref(fig:gems-training)).

(ref:gems-training) Results of training procedure. Participants were randomly
assigned to either the orientation or the spatial frequency condition.

```{r gems-training, fig.width=10, fig.height=4, fig.cap="(ref:gems-training)"}
gridExtra::grid.arrange(
  training_positions_plot + ggtitle("A. Position"),
  training_distance_plot + ggtitle("B. Distance"),
  training_scores_plot + ggtitle("C. Score"),
  nrow = 1
)
```

(ref:gems-training-sensitivities) Sensitivity to stimulus dimensions.

```{r gems-training-sensitivities, fig.width=8, fig.height=3, fig.cap="(ref:gems-training-sensitivities)"}
gridExtra::grid.arrange(
  training_sensitivity_ori_plot +
    ggtitle("A. Sensitivity to orientation"),
  training_sensitivity_sf_plot +
    ggtitle("B. Sensitivity to spatial frequency"),
  nrow = 1
)
```

## First generation {-}

(ref:gen1-results) Results of first generation problem solvers. Each line is a
participant. **A. Position** on the solution landscape. **B. Distance** between current location and 2D peak. **C. Score** over trials.

```{r gen1-results, fig.width=8, fig.height=10, fig.cap="(ref:gen1-results)"}
gridExtra::grid.arrange(
  gen1_positions_plot + ggtitle("A. Position"),
  gen1_distance_plot + ggtitle("B. Distance"),
  gen1_scores_plot + ggtitle("C. Score"),
  nrow = 3
)
```

(ref:gen1-final-scores) First generation performance by block.

```{r gen1-final-scores, fig.width=6, fig.cap="(ref:gen1-final-scores)"}
gridExtra::grid.arrange(
  gen1_final_scores_plot + ggtitle("A. Final scores"),
  gen1_final_distances_plot + ggtitle("B. Distance to peak"),
  nrow = 1
)
```

## Second generation {-}

(ref:gen2-results) Second generation results.

```{r gen2-results, fig.width=8, fig.height=10, fig.cap="(ref:gen2-results)"}
gridExtra::grid.arrange(
  gen2_positions_plot + ggtitle("A. Position"),
  gen2_distance_plot + ggtitle("B. Distance"),
  gen2_scores_plot + ggtitle("C. Score"),
  nrow = 3
)
```

(ref:gen2-final-scores) Second generation performance by block.

```{r gen2-final-scores, fig.width=6, fig.cap="(ref:gen2-final-scores)"}
gridExtra::grid.arrange(
  gen2_final_scores_plot + ggtitle("A. Final scores"),
  gen2_final_distances_plot + ggtitle("B. Distance to peak"),
  nrow = 1
)
```

## Strategies {-}

```{r gems-generation2, fig.width=8}
Gems %>%
  filter(subj_id == "GEMS120" | ancestor_id == "GEMS120",
         landscape_ix != 0, team_trial <= 40) %>%
  ggplot() +
    aes(team_trial, score, group = subj_id, color = instructions) +
    geom_line(aes(linetype = factor(generation))) +
    facet_wrap("landscape_ix", nrow = 1) +
    t_$theme +
    theme(legend.position = "bottom")
```

(ref:strategies) Hill climbing performance. **A. Scores** correspond to heights
on the landscape, and **C. Distances** correspond to Euclidean distances between
current location and peak in the landscape. **B.** and **D.** show these same
measures controlling for starting position.

```{r strategies, fig.width=7, fig.height=7, fig.cap="(ref:strategies)"}
gridExtra::grid.arrange(
  strategies_scores_plot + ggtitle("A. Final score"),
  strategies_relative_scores_plot + ggtitle("B. Relative score"),
  strategies_distance_plot + ggtitle("C. Distance to peak"),
  strategies_relative_distance_plot + ggtitle("D. Relative distance"),
  nrow = 2
)
```

# Discussion {-}

# Appendix {-}

(ref:training) Exact instructions given to participants in either training condition. The instructions differ in the first paragraph and in the sample gabors shown
on the screen.

```{r training, fig.width=8, fig.cap="(ref:training)"}
gridExtra::grid.arrange(
  crotchet::read_image("training_orientation", package = "gems"),
  crotchet::read_image("training_spatial_frequency", package = "gems"),
  nrow = 1
)
```

(ref:training-trial) Training trial procedure. Participants received
feedback in the form of relative changes in their score for each of
the six stimuli on a trial.

```{r training-trial, fig.width=8, fig.cap="(ref:training-trial)"}
gridExtra::grid.arrange(
  crotchet::read_image("training_trial", package = "gems"),
  crotchet::read_image("training_trial_feedback", package = "gems"),
  nrow = 1
)
```

(ref:test-trial) Test trial procedure. Participants received feedback in the
form of the relative change in their score after each selection.

```{r test-trial, fig.width=8, fig.cap="(ref:test-trial)"}
gridExtra::grid.arrange(
  crotchet::read_image("test_trial", package = "gems"),
  crotchet::read_image("test_trial_feedback", package = "gems"),
  nrow = 1
)
```
