---
title: Teams of complementary problem solvers outperform equally skilled individuals
bibliography: references.bib
output:
  bookdown::pdf_document2:
    toc: false
  bookdown::html_document2: {}
---
```{r inherited-diversity-config, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE, results="hide",
               fig.height=4, fig.width=4)
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "inherited-diversity.R"))
```
```{r inherited-diversity}
```

# Introduction {-}

One of the fundamental processes in cultural evolution is the transmission of
problem solving knowledge from one generation to the next. Cultural
transmission is notoriously imperfect when compared to genetic transmission.
Yet the human ability to transmit cultural information far exceeds the
abilities of any non-human animals. Humans have evolved a number of abilities
that improve the success of cultural transmission, including teaching,
language, and other social learning abilities, such as the ability to learn
through observation. A consequence of our system of cultural inheritance is
that problem solvers rarely have to start from scratch when solving new
problems. Instead, they start off with the solutions that were inherited from
a previous generation. A question that arises when considering the consequences
of cultural transmission on human problem solving is the extent to which
future generations are able to benefit from the solutions discovered by previous
generations. The purpose of this research was to understand the conditions
under which inheriting from a previous generation may improve future problem
solving performance beyond what would be achieved by working alone.

Although the human capacity to work together in teams to solve large problems
is evident in many scientific and technological domains, teamwork is rarely
considered to be an effective means of improving individual problem solving
performance. When both teams and individuals complete tasks that individuals
can complete alone, teamwork does not usually improve performance over what
could have been accomplished individually. Part of the reason is due to the
social dynamics of teamwork. Large teams are subject to effects
such as social loafing that decrease motivation and performance. Teamwork
may be inefficient for other reasons, as well. For example, on software
development projects, adding more team members can sometimes slow the speed at
which new changes can be implemented, due to increasing costs of communication
among team members working independently.

A notable exception is a study by @Bahrami:2010jl who found that teams of two
individuals or dyads performed better than individuals in a perceptual judgment
task. In the experiment, participants searched for an oddball target and
responded with a binary judgment corresponding to which of two displays
contained the oddball target. If the dyad disagreed about which display
contained the target, they were allowed to confer and make a joint decision.
Because participants made both individual and team responses on each trial, the
researchers were able to compare the performance of the team to the performance
of best individual in each team. The results showed a clear benefit to teamwork
over what would have been achieved by the best individual in each dyad
completing the task alone. Additional studies have shown that dyads were able
to improve team performance through confidence matching, which is improved by
settling on a common vocabulary to use to communicate uncertainty, but
confidence matching also occurs when using non-linguistic indicators of
uncertainty.

In confidence matching tasks, teamwork improves performance specifically when
the members of the team have similar perceptual abilities. This finding
contrasts with other research on teamwork that has found a benefit to diversity
among team members. In tasks requiring problem solvers to search a space of
possible solutions, homogeneous teams of problem solvers, where all team members
have equal skill, may prematurely convergence on suboptimal solutions, whereas
heterogeneous teams are able to avoid these pitfalls, and maintain a wider
search for globally optimal solutions. These results have been used to argue
that diversity among team members is an effective means of improving team
performance.

To answer this question, we had participants complete a problem solving task
either as first or second generation problem solvers. First generation problem
solvers attempted a problem solving task that simulated a hill climbing
procedure. All first generation problem solvers started off from the same
Second generation problem solvers attempted the same problem solving
task but starting at where the first generation problem solvers had been
at the halfway point.

# Methods {-}

```{r methods}
```

Participants completed a perceptual discrimination task where they were
presented displays of sinusoidal gabor patch stimuli that varied in orientation
and spatial frequency. We asked them to think of the gabors as precious gems on a
foreign planet. Their task was to learn which gems were more valuable than
others.

The value of each gem was determined by its orientation and spatial frequency.
The space of possible gems and their associated values can be visualized as a
three dimensional landscape. As participants selected gems across trials, they
effectively explored different parts of this landscape, looking for the highest
peaks, corresponding to the most valuable gems. In order to select the most
valuable gems on each trial, participants had to learn which stimulus dimensions
were the most important.

On each trial, participants selected one of six gabor patches they thought was
the most valuable. The six gabors presented on each trial were sampled within a
certain distance in the stimulus space of the previously selected gem. In order
to find the most valuable gems, participants had to traverse the solution
landscape in a series of incremental steps.

Participants were trained to pay attention to either the orientation or the bar
width of the gems. The training procedure consisted of written instructions
and 30 training trials. The training procedure is detailed below.

Participants were also assigned to the first or second generation. First
generation participants started at the origin `(0,0)` on each landscape. Second
generation participants were randomly assigned to inherit from a first
generation participant. They inherited the location on the landscape of the gem
selected by the first generation participant after a certain number of trials.
There was no interaction between first and second generation participants other
than that second generation participants started each landscape off in a
different part of the landscape than their first generation predecessors.

In our analyses, we compared three problem solving strategies. Congruent teams
were made of participants trained to pay attention to the same stimulus
dimension---either orientation or bar width. Complementary teams were made of
participants trained on different stimulus dimensions. As a control, we also
compared the performance of individuals trained on a single dimension who
completed the same total number of trials on their own. We hypothesized that
complementary teams of problem solvers would be more efficient at traversing
the solution landscapes than congruent teams or isolated individuals trained
to discriminate the same stimulus dimensions.

(ref:gems-team-structures-instructions) Team structures by instructions.
Participants were trained to pay attention to either orientation or bar width,
and assigned to generations. First generation participants completed the task as
isolated individuals. Second generation participants were assigned to inherit
from a first generation participant.

```{r gems-team-structures, fig.width=8, fig.cap="(ref:gems-team-structures-instructions)"}
gridExtra::grid.arrange(
  crotchet::read_graphviz("team-structures-instructions", package = "gems"),
  crotchet::read_graphviz("team-structures-generations", package = "gems"),
  nrow = 1, widths = c(0.45, 0.55)
)
```

## Materials {-}

### Gabors {-}

We created a $71 \times 71$ grid of `r methods$n_gabors_in_landscape` gabor
patches that varied in orientation and spatial frequency. Orientations were sampled at regular
intervals from `r methods$min_ori`° to `r methods$max_ori`°. Spatial frequencies
were sampled at geometric intervals from `r methods$min_sf` cycles/pixel to
`r methods$max_sf` cycles/pixel. A geometric progression was used because high
spatial frequencies are harder to distinguish than low spatial frequencies.

### Solution landscapes {-}

The three different solution landscapes we used are show in Fig.
\@ref(fig:landscapes). There were two training landscapes, and one test
landscape. The training landscapes were constructed so that the value of a
gem was determined completely by one dimension (either orientation or spatial
frequency) and not influenced by the other dimension. For these landscapes, the
peak value was set at 50 units in either stimulus dimension.

Values in the test landscape were determined by both orientation and spatial
frequency, corresponding to a simple hill landscape with a single peak at `(50,50)`.
The simple hill landscape was symmetric with regard to training condition,
such that both instructions conditions were equally informative about the values
of the gems in the test landscape.

(ref:landscapes) Training and test landscapes. Participants assigned to the
orientation or bar width condition completed the training trials on the
corresponding landscape. All participants completed the test trials on the same
simple hill landscape.

```{r landscapes, fig.width=8, fig.cap="(ref:landscapes)"}
gridExtra::grid.arrange(
  orientation_bias_landscape,
  spatial_frequency_bias_landscape,
  simple_hill_landscape,
  nrow = 1
)
```

### Post experiment survey {-}

After finishing the main experiment, participants completed a short survey. We
asked the participants to recall the instructions they received, to describe how
they approached the experiment, and to rate the extent to which they thought the
gems and scores were random. We were particularly interested in if the
participants felt they were moving around space of possible gabors of their own
volition.

## Procedure {-}

Participants were randomly assigned to training condition (orientation, bar
width) and generation (first, second). The researcher told participants
they would be playing a game where they played as a space explorer who is trying
to find valuable gems on a foreign planet. Their goal was to learn to
distinguish the valuable gems from the cheap ones. Then participants read the
instructions corresponding to their assigned training condition, and completed a
block of 30 training trials. After the training trials, they completed four
blocks of 40 test trials. When finished, participants completed the
post-experiment survey. The experiment took around 30 minutes to complete.

### Training {-}

Participants were randomly assigned to one of two training conditions. In the
orientation condition, participants read in their instructions that the value
of each gem was determined by the orientation of the stripes. In the bar width
condition, they read that the value was determined by the width of the bars.

After reading the instructions, participants completed a block of 30 training
trials. Across the 30 trials, participants traversed a landscape corresponding
to their instructions condition. For example, a participant who was instructed
to pay attention to bar width completed the training trials on the landscape
that was biased so that the values of the gems were dependent completely on
spatial frequency and not on orientation.

On the first training trial, participants were shown six gems sampled from
around the origin `(0,0)` of the stimulus space. On subsequent training trials,
gems were sampled from around the position of the gem
$(x_{orientation},y_{bar\ width})$ selected on the previous trial.

In the training trials, after selecting a gem, participants were shown the
values of all six gems, and had to click on the most valuable gem in the display
to continue. This second response was to encourage participants to learn which
gabors were the most valuable so they could select the most valuable gem on each
trial. If the participant had correctly selected the most valuable gem in their
first response, they were congratulated for making the correct choice, and told
to click on the gem they selected again.

### Test {-}

After completing the training trials, participants completed four blocks of 40
test trials. Participants were given a break after each block, and told that for
the next block, they may encounter gems that they had not seen before.

All test trials were conducted on the same simple hill landscape, defined above.
Participants traversed the simple hill landscape four times. First generation
participants started each block with 6 gems sampled from around the origin
`(0,0)`. Second generation participants participants started off at the location of
a first generation participant after 20 trials. First and second generation
participants were yoked over blocks such that second generation participants
inherited from the same first generation participant a total of four times, one
for each block.

## Participants {-}

Participants were undergraduates at the University of Wisconsin-Madison who
completed the experiment in exchange for course credit. We decided to collect
data for 40 first generation participants, and 80 second generation
participants. There were twice as many second generation participants because
each first generation participant was yoked to two second generation
participants, one in each training condition.

# Results {-}

```{r results}
```

## Training {-}

The purpose of the training procedure was to induce participants to pay
attention to independent dimensions of the gabor stimuli. To verify the success
of the training procedure, we measured the extent to which participants
selected stimuli in each training trial corresponding to their training
condition.

Participants in both training conditions traversed the same stimulus space
in orthogonal directions (Fig. \@ref(fig:gems-training)). Participants selected
gems that moved them closer to the most valuable gems on their particular
landscape. The relationship was linear,
`r r_$training_distance_linear_trial`, with most participants reaching the peak
of the training landscape within 30 trials.

Although participants in both training conditions proceeded in orthogonal
directions, they did not do so equally. Participants trained on bar width were
able to move further toward their peak gem in fewer trials than participants
trained on orientation, `r r_$training_distance_linear_trial_v_instructions`.
This difference between training conditions was unintentional, but it does not
challenge the goals of the experiment.

(ref:gems-training) Results of training procedure. Participants were randomly
assigned to either the orientation or the bar width condition.
**A. Position.** Participants in either training condition traversed the
same stimulus space in orthogonal directions, each progressing toward the
optimal stimulus value in the trained dimension. **B. Distance.** Grid positions
were converted to distances relative to the optimal value in each dimension.
**C. Score.** Cumulative increases in score over trials.

```{r gems-training, fig.width=10, fig.height=4, fig.cap="(ref:gems-training)"}
gridExtra::grid.arrange(
  training_positions_plot + ggtitle("A. Position"),
  training_distance_plot + ggtitle("B. Distance"),
  training_scores_plot + ggtitle("C. Score"),
  nrow = 1
)
```

We also measured the perceptual sensitivity of participants to the different
stimulus dimensions on a trial-by-trial basis. For this analysis, we transformed
the stimulus dimensions of the gems displayed on each trial to be relative to
the optimal value along that dimension. For example, for a participant assigned
to the orientation training condition, gems with orientations at at grid points
40 and 60 were both 10 units away from the peak orientation at 50 units. These
relative values were then mean centered for each trial such that positive values
indicate that the gem was closer to the optimal gem in that dimension than the
other gems on that trial. Using these tranformed values, we fit two models, one
predicting sensitivity to the orientation dimension, and another predicting
sensitivity to the bar width dimension. In the models, we included the
instruction condition of the participants. We expected participants in either
training condition to be more sensitive to to their trained dimension than the
untrained dimension.

As expected, participants trained on either stimulus dimension selected
gems on each trial that moved them closer to the optimal stimulus value
in that dimension, orientation: `r r_$training_sensitivity_ori`, bar width: `r r_$training_sensitivity_sf`.

(ref:gems-training-sensitivities) Sensitivity to stimulus dimensions in the training
procedure.

```{r gems-training-sensitivities, fig.width=6, fig.height=3, fig.cap="(ref:gems-training-sensitivities)"}
gridExtra::grid.arrange(
  training_sensitivity_plot,
  training_sensitivity_trained_dimensions_plot,
  nrow = 1,
  widths = c(0.66, 0.34)
)
```

## First generation {-}

(ref:gen1-results) Results of first generation problem solvers. Each line is a
participant. **A. Position** on the solution landscape. **B. Distance** between current location and 2D peak. **C. Score** over trials.

```{r gen1-results, fig.width=8, fig.height=10, fig.cap="(ref:gen1-results)"}
gridExtra::grid.arrange(
  gen1_positions_plot + ggtitle("A. Position"),
  gen1_distance_plot + ggtitle("B. Distance"),
  gen1_scores_plot + ggtitle("C. Score"),
  nrow = 3
)
```

(ref:gen1-final-scores) First generation performance by block.

```{r gen1-final-scores, fig.width=6, fig.cap="(ref:gen1-final-scores)"}
gridExtra::grid.arrange(
  gen1_final_scores_plot + ggtitle("A. Final scores"),
  gen1_final_distances_plot + ggtitle("B. Distance to peak"),
  nrow = 1
)
```

## Second generation {-}

(ref:gen2-results) Second generation results.

```{r gen2-results, fig.width=8, fig.height=10, fig.cap="(ref:gen2-results)"}
gridExtra::grid.arrange(
  gen2_positions_plot + ggtitle("A. Position"),
  gen2_distance_plot + ggtitle("B. Distance"),
  gen2_scores_plot + ggtitle("C. Score"),
  nrow = 3
)
```

(ref:gen2-final-scores) Second generation performance by block.

```{r gen2-final-scores, fig.width=6, fig.cap="(ref:gen2-final-scores)"}
gridExtra::grid.arrange(
  gen2_final_scores_plot + ggtitle("A. Final scores"),
  gen2_final_distances_plot + ggtitle("B. Distance to peak"),
  nrow = 1
)
```

## Strategies {-}

```{r gems-generation2, fig.width=8}
Gems %>%
  filter(subj_id == "GEMS120" | ancestor_id == "GEMS120",
         landscape_ix != 0, team_trial <= 40) %>%
  ggplot() +
    aes(team_trial, score, group = subj_id, color = instructions) +
    geom_line(aes(linetype = factor(generation))) +
    facet_wrap("landscape_ix", nrow = 1) +
    t_$theme +
    theme(legend.position = "bottom")
```

(ref:strategies) Hill climbing performance. **A. Scores** correspond to heights
on the landscape, and **C. Distances** correspond to Euclidean distances between
current location and peak in the landscape. **B.** and **D.** show these same
measures controlling for starting position.

```{r strategies, fig.width=7, fig.height=7, fig.cap="(ref:strategies)"}
gridExtra::grid.arrange(
  strategies_scores_plot + ggtitle("A. Final score"),
  strategies_relative_scores_plot + ggtitle("B. Relative score"),
  strategies_distance_plot + ggtitle("C. Distance to peak"),
  strategies_relative_distance_plot + ggtitle("D. Relative distance"),
  nrow = 2
)
```

## Post experiment survey {-}

```{r survey}
data("Survey")

questions <- data_frame(question_text = colnames(Survey)) %>%
  mutate(question_id = c("timestamp", "subj_id", "computer", "age", "gender", paste0("Q", 1:9))) %>%
  select(question_id, question_text)

get_question_text <- function(question_id) {
  questions[questions$question_id == question_id, ]$question_text
}

colnames(Survey) <- questions$question_id

predictable_plot <- ggplot(Survey) +
  aes(Q7) +
  geom_bar(stat = "count", width = 0.95) +
  scale_x_continuous("", breaks = 1:5, labels = c("1\nRandom", 2, 3, 4, "5\nPredictable")) +
  ggtitle(get_question_text("Q7")) +
  coord_cartesian(expand = FALSE) +
  t_$theme +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```

# Discussion {-}




