---
title: Teams of complementary problem solvers outperform equally skilled individuals
bibliography: references.bib
output:
  bookdown::pdf_document2:
    toc: false
  bookdown::html_document2: {}
---
```{r inherited-diversity-config, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE, results="hide",
               fig.height=4, fig.width=4)
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "inherited-diversity.R"))

library(crotchet)
read_graphviz_chunk("team-structures-instructions", package = "gems",
                    new_name = "gems-team-structures-instructions")
read_graphviz_chunk("team-structures-generations", package = "gems",
                    new_name = "gems-team-structures-generations")
```
```{r inherited-diversity}
```

# Introduction {-}

One of the fundamental processes in cultural evolution is the transmission of
tools and other solutions to problems from one generation to the next. Cultural
transmission is notoriously imperfect when compared to genetic transmission.
Yet the human ability to transmit cultural information far exceeds the
abilities of any non-human animals. Humans have evolved a number of abilities
that improve the success of cultural transmission, including teaching,
language, and other social learning abilities, such as the ability to learn
through observation. A consequence of our system of cultural inheritance is
that problem solvers rarely have to start from scratch when solving new
problems. Instead, they start off with the solutions that were inherited from
a previous generation. The purpose of this research was to understand how
inheriting from a previous generation influences future problem solving
behavior.

At the population level, our advanced social learning abilities enable certain
complex traits to be sustained over generations even if the

Cultural transmission across generations is notoriously imperfect. For example,
when folktales are passed from person to person in a transmission chain
procedure, errors and omissions are introduced into the stories, and they
slowly lose their resemblance to the original story.

However, a number of experiments have demonstrated that the information that is
lost or retained through repeated storytelling is not random. Bartlett's
original demonstrations of this phenomena show that culturally learned schema
for how folktales are structured influences which information is retained over
generations. More recent investigations have found that social information, as
opposed to factual information, is more likely to be retained over generations
[@Mesoudi].

In some cases, the loss of information over generations is beneficial. Flynn
[-@Flynn] found that as children repeatedly copied the same actions over
generations, the actions became simpler, and lost their resemblance to the
original action.

Some starting points are more beneficial than others. A laptop is clearly more
beneficial than a calculator. But does the starting point affect the way in
which the problem is solved? If the starting point can improve future problem
solving performance, then our system of cultural inheritance may actually
contribute to a general improvement in human problem solving ability.

To answer this question, we had participants complete a problem solving task
either as first or second generation problem solvers. First generation problem
solvers attempted a problem solving task that simulated a hill climbing
procedure. All first generation problem solvers started off from the same
Second generation problem solvers attempted the same problem solving
task but starting at where the first generation problem solvers had been
at the halfway point.

A consequence of our system of cultural inheritance is that problem solvers
rarely have to start from scratch when solving new problems. Instead, they start
off with the solutions that were inherited from a previous generation.

There are many ways to measure the effectiveness of teamwork. A common method
is to compare the performance of teams to individuals working alone. When
compared in this way, studies have often failed to find any evidence that
teamwork provides an improvement over individual problem solving.

One of the reasons teamwork is reflected rather poorly in the experimental
literature is due to limits in experimental methods. In order to conduct an
experiment in which both teams and individuals are administered the same task,
many tasks are necessarily excluded because they can not be equally attempted
using both teams and individuals. To use a clear example: if the problem
solving task required moving heavy objects, teamwork has clear advantages, but
no one would be surprised by these results.

Teams can clearly solve many problems that are insurmountable for individuals.
But when both teams and individuals attempt tasks that individuals can complete
alone, teamwork is rarely found to be an improvement over individual
performance. A notable exception is a study by @Bahrami:2010jl who found that
dyads were better than individuals in a perceptual judgment task. In the
experiment, participants searched for an oddball target and responded with
a binary judgment corresponding to which of two displays contained the target.
If the dyad disagreed about which display contained the target, they were
allowed to confer and make a joint decision. Because the participants made both
individual and team responses on each trial, the researchers were able to
compare the performance of the team to the performance of best individual in
each team. The results showed a clear benefit to teamwork over what would have
been achieved by the best individual in each dyad completing the task alone.

Later studies have shown that dyads were able to improve team performance
through confidence matching, which is improved by settling on a common vocabulary
to use to communicate uncertainty, but confidence matching also occurs
when using continuous measures of uncertainty. In confidence matching tasks,
teamwork improves team performance over what could be achieved by the best
individual specifically when the members of the team have the same perceptual
ability.

A difficulty that arises when comparing experiments done on teamwork
is in how to properly control for time. The typical approach is to control
for **calendar hours**. For example, teams and individuals are both given
an hour to complete the same task. However, controlling for calendar hours does
not scale with team size. Individuals and teams are both given an hour,
regardless of the size of the team.

In the experiment, we compare three strategies for solving the same problem.

# Methods {-}

```{r methods}
```

Participants completed a perceptual discrimination task where they were
presented displays of sinusoidal gabor patch stimuli that varied in orientation and
spatial frequency. We asked them to think of the gabors as "gems"
from a foreign planet. Their task was to learn which gems were
more valuable than others.

The value of each gem was determined by its orientation and spatial frequency.
The space of possible gems and their associated values can be visualized as a
three dimensional landscape. As participants selected gems across trials, they
effectively explored different parts of this landscape, looking for the highest
peaks, corresponding to the most valuable gems. In order to detect
the most valuable gems on each trial, participants had to learn which stimulus
dimensions were the most important.

On each trial, participants selected one of six gabor patches they thought
was the most valuable. The six gabors presented on each trial were sampled
within a certain distance of previously selected gem. In order to find the most
valuable gems, participants had to navigate the solution landscape in a series
of incremental steps.

Participants were trained to pay attention to either the orientation or the bar
width of the gems. The training procedure consisted of written instructions
and 30 training trials. The training procedure is detailed below.

Participants were also assigned to the first or second generation. First
generation participants started at the origin `(0,0)` on each landscape. Second
generation participants were randomly assigned to inherit from a first
generation participant. They inherited the location on the landscape of the gem
selected by the first generation participant. There was no interaction between
first and second generation participants other than that second generation
participants started each landscape off in a different part of the landscape
than their predecessors.

In our analyses, we compare three problem solving strategies. Congruent teams
were made of participants trained to pay attention to the same stimulus
dimension---either orientation or bar width. Complementary teams were made of
participants trained on different stimulus dimensions. As a control, we also
compared the performance of individuals trained on a single dimension who
completed the same total number of trials on their own.

(ref:gems-team-structures-instructions) Team structures by instructions.
Participants were trained to pay attention to either orientation or bar width,
and assigned to generations. First generation participants completed the task as
isolated individuals. Second generation participants were assigned to inherit
from a first generation participant.

```{r gems-team-structures, fig.width=8, fig.cap="(ref:gems-team-structures-instructions)"}
gridExtra::grid.arrange(
  crotchet::read_graphviz("team-structures-instructions", package = "gems"),
  crotchet::read_graphviz("team-structures-generations", package = "gems"),
  nrow = 1, widths = c(0.45, 0.55)
)
```

## Materials {-}

### Gabors {-}

We created a 71x71 grid of `r methods$n_gabors_in_landscape` gabor patches that
varied in orientation and spatial frequency. A sample of the space is shown in
Fig. \@ref(fig:gabors). Orientations were sampled at regular intervals from
`r methods$min_ori`° to `r methods$max_ori`°. Spatial frequencies were sampled at
geometric intervals from `r methods$min_sf` cycles/pixel to `r methods$max_sf`
cycles/pixel. A geometric progression was used because high spatial frequencies
are harder to distinguish than low spatial frequencies.

(ref:gabors) Sampled space of possible gabor patches. Gabor patches varied
in orientation and spatial frequency.

```{r gabors, fig.cap="(ref:gabors)"}
draw_image("SimpleHillGems", package = "gems")
```

### Solution landscapes {-}

The three different solution landscapes we used are show in Fig.
\@ref(fig:landscapes). There were two training
landscapes, and one test landscape. The training landscapes were constructed so
that the value of a gem was determined completely by one dimension (either
orientation or spatial frequency) and not influenced by the other dimension. For
these landscapes, the peak value was set at 50 units in either stimulus dimension.

Values in the test landscape were determined by both orientation and spatial
frequency, corresponding to a simple hill landscape with a single peak at `(50,50)`.
The simple hill landscape was symmetric with regard to training condition,
such that both instructions conditions were equally helpful for the test
landscape.

(ref:landscapes) Training and test landscapes. Participants assigned to the
orientation or bar width condition completed the training trials on the
corresponding landscape. All participants completed the test trials on the same
simple hill landscape.

```{r landscapes, fig.width=8, fig.cap="(ref:landscapes)"}
gridExtra::grid.arrange(
  orientation_bias_landscape,
  spatial_frequency_bias_landscape,
  simple_hill_landscape,
  nrow = 1
)
```

### Post experiment survey {-}

After finishing the main experiment, participants completed a short survey.
We asked the participants about the instructions they received, how they
approached the experiment, and to what extent they thought the gems and
scores were random. We were particularly interested in if the
participants felt they were moving around space of possible gabors
of their own volition. The questions are listed in Table \@ref(tab:survey-questions).

(ref:survey-questions) Questions asked in the post-experiment survey.

```{r survey-questions, results='asis'}
data("Survey")
data_frame(question_text = colnames(Survey)) %>%
  mutate(question_id = c("timestamp", "subj_id", "computer", "age", "gender", paste0("Q", 1:9))) %>%
  filter(grepl("Q", question_id)) %>%
  select(`#` = question_id, `Text` = question_text) %>%
  kable(caption="(ref:survey-questions)") %>%
  column_spec(2, width = "45em")
```

## Procedure {-}

Participants were randomly assigned training condition and generation.
The researcher told all participants that they would be playing a
game where they played as a space explorer who is trying to find valuable
gems on a foreign planet. The goal of the game was to learn to find
the most valuable gems. Then participants read the instructions corresponding
to training condition, and completed a block of 30 training trials. After
the training trials, they completed four blocks of 40 test trials. When
finished, participants completed the post-experiment survey.
The experiment took around 30 minutes to complete.

### Training {-}

Participants were randomly assigned to one of two training conditions. In the
orientation condition, participants read in their instructions that the value
of each gem was determined by the orientation of the stripes. In the bar width
condition, they read that the value was determined by the width of the bars.
The exact instructions are shown in Fig. \@ref(fig:training).

After reading the instructions, participants completed a block of 30 training
trials. Across the 30 trials, participants traversed a landscape corresponding
to their instructions condition. For example, a participant who was instructed
to pay attention to bar width completed the training trials on the landscape
that was biased so that the values of the gems were dependent completely on
spatial frequency and not on orientation.

On the first training trial, participants were shown six gems sampled from around
the origin `(0,0)` of the stimulus space. On subsequent training trials, gems
were sampled from around the position of the gem $(x_{ori},y_{bw})$ selected on the
previous trial.

After selecting a gem, participants were shown the
values of all six gems, and had to click on the most valuable gem in the display
to continue. This second response was to encourage
participants to learn which gabors were the most valuable so they could select
the most valuable gem on each trial. If the participant had correctly selected
the most valuable gem in their first response, they were congratulated for
making the correct choice, and told to click on the gem they selected again. An
example training trial is shown in Fig. \@ref(fig:training-trial).

### Test {-}

After completing the training trials, participants completed four blocks of 40
trials. Participants were given a break after each block, and told that for
the next block, they may encounter gems that they had not seen before.

All test trials were conducted on the same simple hill landscape, defined above.
Participants traversed the simple hill landscape four times. First generation
participants started each block with 6 gems sampled from around the origin
`(0,0)`. Second generation participants participants started off at the location of
a first generation participant after 20 trials. First and second generation
participants were yoked over blocks such that second generation participants
inherited from the same first generation participant a total of four times, one
for each block.

## Participants {-}

Participants were undergraduates at the University of Wisconsin-Madison who
completed the experiment in exchange for course credit. We decided to collect
data for 40 first generation participants, and 80 second generation
participants. There were twice as many second generation participants because
each first generation participant was yoked to two second generation
participants, one in each training condition.

# Results {-}

```{r results}
```

## Training {-}

The purpose of the training procedure was for participants to learn to pay
attention to orthogonal dimensions of the gem stimuli. To verify whether the
training procedure worked, we measured the extent to which participants
selected stimuli in each training trial corresponding to their training
condition.

Participants in either training condition traversed the same stimulus space
in orthogonal directions (Fig. \@ref(fig:gems-training)). Participants selected
gems that moved them closer to the most valuable gems on their particular
landscape. The relationship was linear,
`r r_$training_distance_linear_trial`, with most participants reaching the peak
of the training landscape within 30 trials.

Although participants in both training conditions proceeded in orthogonal
directions, they did not do so equally. Participants trained on bar width were
able to move further toward their peak gem in fewer trials than participants
trained on orientation, `r r_$training_distance_linear_trial_v_instructions`.

(ref:gems-training) Results of training procedure. Participants were randomly
assigned to either the orientation or the spatial frequency condition.
**A. Position.** Participants in either training condition traversed the
same stimulus space in orthogonal directions, each progressing toward the
optimal stimulus value in the trained dimension. **B. Distance.** Grid positions
were converted to distances relative to the optimal value in each dimension.
**C. Score.** Cumulative increases in score over trials.

```{r gems-training, fig.width=10, fig.height=4, fig.cap="(ref:gems-training)"}
gridExtra::grid.arrange(
  training_positions_plot + ggtitle("A. Position"),
  training_distance_plot + ggtitle("B. Distance"),
  training_scores_plot + ggtitle("C. Score"),
  nrow = 1
)
```

We also measured the perceptual sensitivity of participants to the different
stimulus dimensions. For our analysis, we transformed the stimulus dimensions
of the gems displayed on each trial to be relative to the optimal value along
that dimension. These relative values were then mean centered so that they
formed a normal distribution. Positive values in the transformed stimulus
dimension mean that the gem that was selected on a particular trial was closer
to the optimal gem in that dimension. We fit two models, one predicting
sensitivity to the orientation dimension, and one predicting
sensitivity to the spatial frequency dimension. In the models, we included
the instruction condition of the participants. We expected participants
in either training condition to be more sensitive to to their trained dimension
than the orthogonal dimension.

(ref:gems-training-sensitivities) Sensitivity to stimulus dimensions.

```{r gems-training-sensitivities, fig.width=6, fig.height=3, fig.cap="(ref:gems-training-sensitivities)"}
training_sensitivity_plot
```

## First generation {-}

(ref:gen1-results) Results of first generation problem solvers. Each line is a
participant. **A. Position** on the solution landscape. **B. Distance** between current location and 2D peak. **C. Score** over trials.

```{r gen1-results, fig.width=8, fig.height=10, fig.cap="(ref:gen1-results)"}
gridExtra::grid.arrange(
  gen1_positions_plot + ggtitle("A. Position"),
  gen1_distance_plot + ggtitle("B. Distance"),
  gen1_scores_plot + ggtitle("C. Score"),
  nrow = 3
)
```

(ref:gen1-final-scores) First generation performance by block.

```{r gen1-final-scores, fig.width=6, fig.cap="(ref:gen1-final-scores)"}
gridExtra::grid.arrange(
  gen1_final_scores_plot + ggtitle("A. Final scores"),
  gen1_final_distances_plot + ggtitle("B. Distance to peak"),
  nrow = 1
)
```

## Second generation {-}

(ref:gen2-results) Second generation results.

```{r gen2-results, fig.width=8, fig.height=10, fig.cap="(ref:gen2-results)"}
gridExtra::grid.arrange(
  gen2_positions_plot + ggtitle("A. Position"),
  gen2_distance_plot + ggtitle("B. Distance"),
  gen2_scores_plot + ggtitle("C. Score"),
  nrow = 3
)
```

(ref:gen2-final-scores) Second generation performance by block.

```{r gen2-final-scores, fig.width=6, fig.cap="(ref:gen2-final-scores)"}
gridExtra::grid.arrange(
  gen2_final_scores_plot + ggtitle("A. Final scores"),
  gen2_final_distances_plot + ggtitle("B. Distance to peak"),
  nrow = 1
)
```

## Strategies {-}

```{r gems-generation2, fig.width=8}
Gems %>%
  filter(subj_id == "GEMS120" | ancestor_id == "GEMS120",
         landscape_ix != 0, team_trial <= 40) %>%
  ggplot() +
    aes(team_trial, score, group = subj_id, color = instructions) +
    geom_line(aes(linetype = factor(generation))) +
    facet_wrap("landscape_ix", nrow = 1) +
    t_$theme +
    theme(legend.position = "bottom")
```

(ref:strategies) Hill climbing performance. **A. Scores** correspond to heights
on the landscape, and **C. Distances** correspond to Euclidean distances between
current location and peak in the landscape. **B.** and **D.** show these same
measures controlling for starting position.

```{r strategies, fig.width=7, fig.height=7, fig.cap="(ref:strategies)"}
gridExtra::grid.arrange(
  strategies_scores_plot + ggtitle("A. Final score"),
  strategies_relative_scores_plot + ggtitle("B. Relative score"),
  strategies_distance_plot + ggtitle("C. Distance to peak"),
  strategies_relative_distance_plot + ggtitle("D. Relative distance"),
  nrow = 2
)
```

## Post experiment survey {-}

```{r survey}
data("Survey")

questions <- data_frame(question_text = colnames(Survey)) %>%
  mutate(question_id = c("timestamp", "subj_id", "computer", "age", "gender", paste0("Q", 1:9))) %>%
  select(question_id, question_text)

get_question_text <- function(question_id) {
  questions[questions$question_id == question_id, ]$question_text
}

colnames(Survey) <- questions$question_id

predictable_plot <- ggplot(Survey) +
  aes(Q7) +
  geom_bar(stat = "count", width = 0.95) +
  scale_x_continuous("", breaks = 1:5, labels = c("1\nRandom", 2, 3, 4, "5\nPredictable")) +
  ggtitle(get_question_text("Q7")) +
  coord_cartesian(expand = FALSE) +
  t_$theme +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```

# Discussion {-}

# Appendix {-}

(ref:training) Exact instructions given to participants in either training condition. The instructions differ in the first paragraph and in the sample gabors shown
on the screen.

```{r training, fig.width=8, fig.cap="(ref:training)"}
gridExtra::grid.arrange(
  crotchet::read_image("training_orientation", package = "gems"),
  crotchet::read_image("training_spatial_frequency", package = "gems"),
  nrow = 1
)
```

(ref:training-trial) Training trial procedure. Participants received
feedback in the form of relative changes in their score for each of
the six stimuli on a trial.

```{r training-trial, fig.width=8, fig.cap="(ref:training-trial)"}
gridExtra::grid.arrange(
  crotchet::read_image("training_trial", package = "gems"),
  crotchet::read_image("training_trial_feedback", package = "gems"),
  nrow = 1
)
```

(ref:test-trial) Test trial procedure. Participants received feedback in the
form of the relative change in their score after each selection.

```{r test-trial, fig.width=8, fig.cap="(ref:test-trial)"}
gridExtra::grid.arrange(
  crotchet::read_image("test_trial", package = "gems"),
  crotchet::read_image("test_trial_feedback", package = "gems"),
  nrow = 1
)
```
