---
title: Complementary inheritance in iterated problem solving
bibliography: references.bib
output:
  bookdown::pdf_document2:
    toc: false
    includes:
      in_header: templates/header.tex
abstract: |
  Humans have evolved a system of cultural inheritance that enables the
  accumulation of problem solving knowledge over generations. As a result,
  we rarely have to start from scratch when solving new problems, and instead
  can benefit from solutions discovered by a previous generation. We
  investigated the impact of inheritance on problem solving performance.
  Specifically, we explored the conditions under which inheritance might
  improve problem solving relative to solving the same problems individually.
  In Experiment 1, we simulated the effect of inheritance on problem solving
  using agent-based models, and demonstrate some problems for
  which generations of complementary problem solvers, who learn independent and
  complementary skills, are able to outperform equally skilled individuals. In
  Experiment 2, we tested these predictions in a behavioral experiment. These
  results illustrate one way in which our system of cultural inheritance may
  result in cumulative improvements to problem solving performance over
  generations.
---
```{r inherited-diversity-config, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE, results="hide",
               fig.height=4, fig.width=4,
               autodep = TRUE)
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "inherited-diversity.R"))
read_chunk(file.path(chunks_dir, "peaks.R"))
```
```{r inherited-diversity}
```

# Introduction {-}

One of the fundamental processes in cultural evolution is the transmission of
problem solving knowledge from one generation to the next. Cultural
transmission is notoriously imperfect when compared to genetic transmission
[@Sperber:1996exc; @Claidiere:2010kd]. Yet the human ability to transmit
cultural information far exceeds the abilities of any non-human animals
[@Tomasello:1993vh; @Tennie:2009gaa; @Dean:2012gv]. One important consequence
of our system of cultural inheritance is that we rarely have to start from
scratch when solving new problems. Instead, we often start off with solutions
that were inherited from a previous generation. The purpose of this research
was to investigate whether inheriting from a previous generation could ever
improve problem solving beyond what could be achieved individually.

Humans have evolved a number of abilities including imitation, teaching,
and language that improve the success of cultural inheritance. These abilities
enable high fidelity cultural transmission [@Lewis:2012jg; @Dean:2012gv].
That some behaviors can be accurately copied through social learning
is uncontroversial, but what is in active debate is the importance of accurate
transmission in explaining the nature of human cultural evolution. For instance,
Sperber and others [-@Sperber:1996exc; @Claidiere:2010kd] have argued that
accurate copying is possible but rare, and that cultural transmission is better
understood as an inferential and transformative process as opposed to a
replicative one.

Tomasello and others [-@Tomasello:1993vh; @Tennie:2009gaa] have argued that
accurate transmission is clearly important in many problem solving domains.
Accurate transmission of problem solving knowledge is important for two
reasons. First, it allows others to benefit from complex solutions they
themselves could not have created. Second, it increases the likelihood that
future generations cumulatively improve upon the solutions discovered by their
ancestors, even if these improvements are discovered by chance. On this view,
the capacity to incrementally improve solutions to problems inherited from
others is a driving force in human cultural evolution.

The primary benefit to inheriting from a previous generation is that it gives
future generations a head start on problem solving. Rather than each new
generation striving to solve the same problems over and over again, future
generations can save the time it would have taken to solve the same problem on
their own. Instead, they can spend their time solving new problems, not
attempted by previous generations. As a result, cultural inheritance allows for
problem solving to extend far longer than any individual lifetime. Over time,
we have come to depend on this system of cultural inheritance to the point that
human cultures are able to solve problems that no individual could attempt on
their own [@Henrich:2015srt; @Boyd:2017diff].

Importantly, on this view, cultural inheritance improves the problem solving
performance of the group without affecting underlying problem solving ability.
Future generations get a head start on future problem solving, but they are not
fundamentally different types of problem solvers. Here we asked whether some
types of inheritance are more beneficial than others. Specifically we asked
whether generations of problem solvers trained on independent but complementary
skills can outperform individual problem solvers given the same training.

Although the human capacity to work together in teams to solve large problems
is evident in many scientific and technological domains, teamwork is rarely
considered to be an effective means of improving individual problem solving
performance. When both teams and individuals complete tasks that individuals
can complete alone, teamwork does not usually improve performance over what
could have been accomplished individually. Part of the reason is due to the
social dynamics of teamwork. Large teams are subject to effects such as social
loafing that decrease motivation and performance. Teamwork may be inefficient
for other reasons, as well. For example, on software development projects,
adding more team members can sometimes slow the speed at which new changes can
be implemented, due to increasing costs of communication among team members
working independently.

A notable exception is a study by @Bahrami:2010jl who found that teams of two
individuals or dyads performed better than individuals in a perceptual judgment
task. In the experiment, participants searched for an oddball target and
responded with a binary judgment corresponding to which of two displays
contained the oddball target. If the dyad disagreed about which display
contained the target, they were allowed to confer and make a joint decision.
Because participants made both individual and team responses on each trial, the
researchers were able to compare the performance of the team to the performance
of best individual in each team. The results showed a clear benefit to teamwork
over what would have been achieved by the best individual in each dyad
completing the task alone. Additional studies have shown that dyads were able
to improve team performance through confidence matching, which is improved by
settling on a common vocabulary to use to communicate uncertainty, but
confidence matching also occurs when using non-linguistic indicators of
uncertainty.

In confidence matching tasks, teamwork improves performance specifically when
the members of the team have similar perceptual abilities. This finding
contrasts with other research on teamwork that has found a benefit to diversity
among team members. In tasks requiring problem solvers to search a space of
possible solutions, homogeneous teams of problem solvers, where all team members
have equal skill, may prematurely convergence on suboptimal solutions, whereas
heterogeneous teams are able to avoid these pitfalls, and maintain a wider
search for globally optimal solutions. These results have been used to argue
that diversity among team members is an effective means of improving team
performance.

To answer this question, we had participants complete a problem solving task
either as first or second generation problem solvers. First generation problem
solvers attempted a problem solving task that simulated a hill climbing
procedure. All first generation problem solvers started off from the same
Second generation problem solvers attempted the same problem solving
task but starting at where the first generation problem solvers had been
at the halfway point.

We wanted to understand the conditions under which inheritance across
individuals results in performance greater than either individual could achieve
on their own. We hypothesized that as the could result in performance greater
than either individual could achieve on their own.

# Experiment 1 {-}

```{r peaks-setup}
```

In Experiment 1, we used an agent-based model to simulate the effect of
inheritance on problem solving performance. Agents were assigned skills in two
independent dimensions. Some agents had equal skill in both dimensions, while
others had unequal skill, specializing along one dimension at the expense of
skill in the other. Agents with each type of skill distribution were arranged
in pairs with other agents, forming two agent teams that varied in overall
skill distribution. For identical teams, both agents had equal skills along
both dimensions. For diverse teams, each agent specialized in a different
skill. We tested whether skill diversity resulted in better problem solving
performance than equally skilled individuals.

## Methods {-}

```{r peaks-methods}
```

In the simulations, agents completed a problem solving task by searching
a space of possible solutions, looking for the best ones. Agents navigated the
solution landscape using a simple hill climbing procedure. For each time step,
agents randomly sampled a solution within their range of vision (described
below). If the sampled solution corresponded to a higher position on the
landscape than the previous best solution, then the agent moved to the new
position. On the next time step, a solution was sampled relative to the new
position.

Agents were assigned skill values in two dimensions (Fig.
\@ref(fig:ability-as-vision)). Each skill value corresponded to vision along
a particular dimension in the solution landscape. All agents were allotted the
same total amount of skill, which was distributed between the two dimensions.
Some agents had equal skill in both dimensions, and others had unequal skill in
either dimension. Agents with equal skill in both dimensions had a square range
of vision, and could sample equally in any direction. Agents with unequal
skills had a rectangular range of vision, and were more likely to sample in one
dimension over the other.

(ref:ability-as-vision) Methods for agent-based models. **A.** Each agent was
assigned skills in two dimensions. **B.** Skill values in both dimensions were
combined to form each agent's search areas, corresponding to the range of
solutions that could be sampled relative to any point on the landscape. **C.**
Teams of two agents differing in skill diversity. Identical teams shared the
same abilities in both dimensions. Complementary teams had "equal but opposite"
skills.

```{r ability-as-vision, fig.width=8, fig.height=8, fig.cap='(ref:ability-as-vision)'}
grid.arrange(
  gg_two_dimensions + ggtitle("A. Vision in two dimensions"),
  gg_search_rects + ggtitle("B. Search areas"),
  gg_differing_skills + ggtitle("C. Two person teams varying in vision distribution"),
  ncol = 1
)
```

## Results {-}

```{r peaks-experiments}
```

The results of the simulations are shown in Fig. \@ref(fig:differing-skills).

(ref:differing-skills) Results of agent-based model simulations.

```{r differing-skills, fig.width=8, fig.height=4, fig.cap="(ref:differing-skills)"}
limits <- seq(-100, 100, by = 10)
z <- expand.grid(x = limits, y = limits) %>%
     mutate(z = -x^2 - y^2)
simple_hill <- lattice::wireframe(z ~ x * y, data = z)

grid.arrange(simple_hill, gg_differing_skills_timeline, nrow = 1)
```

(ref:differing-skills-walk) Grid positions of agent-based models by skill distribution.

```{r differing-skills-walk, fig.width=8, fig.height=2.5, fig.cap="(ref:differing-skills-walk)"}
gg_differing_skills_walk
```

## Discussion {-}

Experiment 1 demonstrates the process by which inheritance may result in teams of
problem solvers may come to outperform individuals trained on the same skills and
given the same amount of time. In the simulations, this improvement in performance
is the result of the first generation putting later generations in a part of the
problem space that would be unlikely to be achieved individually.

However, the value of the agent-based model is limited, especially since the
problem solving process is simulated as a random walk process. A more realistic
test of whether inheritance can improve team performance requires a task in
which new solutions are tested by participants seeking to answer the same problems.
In Experiment 2, we test the predictions drawn from Experiment 1 in a behavioral
in lab experiment.

# Experiment 2 {-}

In Experiment 2 we tested the predictions from the agent-based model presented
in Experiment 1 in a behavioral experiment. Participants were randomly assigned
to learn equal but independent skills in a perceptual discrimination task. In
the task, participants had to solve a problem by searching a space of possible
solutions looking for the best solutions. We expected to find that teams of
complementary problem solvers, trained on independent skills, could coordinate
their labor so as to improve problem solving performance beyond what either
individual could have achieved on their own.

## Methods {-}

```{r methods}
```

(ref:gems-team-structures) Team structures. Participants were trained to pay
attention to either orientation (ori) or bar width (bw), and assigned to
generations. First generation participants (top row) completed the task as
**isolated** individuals. Second generation participants were assigned to inherit
from a first generation participant trained on the same dimension (**congruent**,
middle row), or on different dimensions (**complementary**, bottom row). Left and
right panels show counterbalanced conditions.

```{r gems-team-structures, fig.width=8, fig.cap="(ref:gems-team-structures)"}
crotchet::draw_graphviz("team-structures-counterbalanced", package = "gems")
```

Participants completed a perceptual discrimination task where they were
presented displays of sinusoidal gabor patch stimuli that varied in orientation
and spatial frequency. We asked them to think of the gabors as precious gems on a
foreign planet. Their task was to learn which gems were more valuable than
others.

The value of each gem was determined by its orientation and spatial frequency.
The space of possible gems and their associated values can be visualized as a
three dimensional landscape. As participants selected gems across trials, they
effectively explored different parts of this landscape, looking for the highest
peaks, corresponding to the most valuable gems. In order to select the most
valuable gems on each trial, participants had to learn which stimulus dimensions
were the most important.

On each trial, participants selected one of six gabor patches they thought was
the most valuable. The six gabors presented on each trial were sampled within a
certain distance from the previously selected gem. The sampling procedure is
described in the procedure, below. As participants selected gems across trials,
they effectively traversed the solution landscape in a series of
incremental steps.

(ref:gems-trial) Procedure. On each trial, participants were shown six gabor patches
referred to as "gems". Each gem was assigned a value. As participants selected gems
across trials, they navigated the stimulus space of all possible gems.

```{r gems-trial, fig.width=6, fig.cap="(ref:gems-trial)"}
gridExtra::grid.arrange(
  crotchet::read_image("trial", package = "gems"),
  trial_plot,
  nrow = 1
)
```

Participants were trained to pay attention to either the orientation or the bar
width of the gems. The training procedure consisted of written instructions
and 30 training trials. The training procedure is detailed below.

Participants were also assigned to the first or second generation. First
generation participants started at the origin $(0_{ori},0_{bw})$ on each
landscape. Second generation participants were assigned to inherit from a first
generation participant. They inherited the location on the landscape
$(x_{ori},y_{bw})$ of the gem selected by the first generation participant after
a certain number of trials. There was no interaction between first and second
generation participants other than that second generation participants started
each landscape off in a different part of the landscape than their first
generation predecessors.

In our analyses, we compared three strategies for finding the most valuable gems
(Fig. \@ref(fig:gems-team-structures)). Congruent teams were made of
participants trained to pay attention to the same stimulus dimension---either
orientation or bar width. Complementary teams were made of participants trained
on different stimulus dimensions. We also compared the performance of
individuals trained on a single dimension who completed the same total number of
trials on their own. We hypothesized that complementary teams of problem solvers
would be more efficient at traversing the solution landscapes than congruent
teams or isolated individuals trained to discriminate the same stimulus
dimension.

### Materials {-}

#### Gabors {-}

We created a $71 \times 71$ grid of `r methods$n_gabors_in_landscape` gabor
patches that varied in orientation and spatial frequency. Orientations were sampled at regular
intervals from `r methods$min_ori`° to `r methods$max_ori`°. Spatial frequencies
were sampled at geometric intervals from `r methods$min_sf` cycles/pixel to
`r methods$max_sf` cycles/pixel. A geometric progression was used because high
spatial frequencies are harder to distinguish than low spatial frequencies.

#### Landscapes {-}

The three different solution landscapes we used are show in Fig.
\@ref(fig:landscapes). There were two training landscapes, and one test
landscape. The training landscapes were constructed so that the value of a
gem was determined completely by one dimension (either orientation or spatial
frequency) and not influenced by the other dimension. For these landscapes, the
peak value was set at 50 units in either stimulus dimension.

Values in the test landscape were determined by both orientation and spatial
frequency, corresponding to a simple hill landscape with a single peak at `(50,50)`.
The simple hill landscape was symmetric with regard to training condition,
such that both instructions conditions were equally informative about the values
of the gems in the test landscape.

(ref:landscapes) Landscapes. Participants assigned to the
orientation or bar width condition completed the training trials on the
corresponding landscape. All participants completed the test trials on the same
simple hill landscape.

```{r landscapes, fig.width=8, fig.cap="(ref:landscapes)"}
gridExtra::grid.arrange(
  orientation_bias_landscape,
  spatial_frequency_bias_landscape,
  simple_hill_landscape,
  nrow = 1
)
```

#### Survey {-}

After finishing the main experiment, participants completed a short survey. We
asked the participants to recall the instructions they received, to describe how
they approached the experiment, and to rate the extent to which they thought the
gems and scores were random. We were particularly interested in if the
participants felt they were moving around space of possible gabors of their own
volition.

### Procedure {-}

Participants were randomly assigned to training condition (orientation, bar
width) and generation (first, second). The researcher told participants
they would be playing a game where they played as a space explorer who is trying
to find valuable gems on a foreign planet. Their goal was to learn to
distinguish the valuable gems from the cheap ones. Then participants read the
instructions corresponding to their assigned training condition, and completed a
block of 30 training trials. After the training trials, they completed four
blocks of 40 test trials. When finished, participants completed the
post-experiment survey. The experiment took around 30 minutes to complete.

#### Training trial procedure {-}

Participants were randomly assigned to one of two training conditions. In the
orientation condition, participants read in their instructions that the value
of each gem was determined by the orientation of the stripes. In the bar width
condition, they read that the value was determined by the width of the bars.

After reading the instructions, participants completed a block of 30 training
trials. Across the 30 trials, participants traversed a landscape corresponding
to their instructions condition. For example, a participant who was instructed
to pay attention to bar width completed the training trials on the landscape
that was biased so that the values of the gems were dependent completely on
spatial frequency and not on orientation.

On the first training trial, participants were shown six gems sampled from
around the origin $(0_{ori},0_{bw})$ of the stimulus space. On subsequent training trials,
gems were sampled from around the position of the gem
$(x_{ori},y_{bw})$ selected on the previous trial.

In the training trials, after selecting a gem, participants were shown the
values of all six gems, and had to click on the most valuable gem in the display
to continue. This second response was to encourage participants to learn which
gabors were the most valuable so they could select the most valuable gem on each
trial. If the participant had correctly selected the most valuable gem in their
first response, they were congratulated for making the correct choice, and told
to click on the gem they selected again.

#### Test trial procedure {-}

After completing the training trials, participants completed four blocks of 40
test trials. Participants were given a break after each block, and told that for
the next block, they may encounter gems that they had not seen before.

All test trials were conducted on the same simple hill landscape, defined above.
Participants traversed the simple hill landscape four times. First generation
participants started each block with 6 gems sampled from around the origin
$(0_{ori},0_{bw})$. Second generation participants participants started off at the location of
a first generation participant after 20 trials, $(x_{ori},y_{bw})$. First and second generation
participants were yoked over blocks such that second generation participants
inherited from the same first generation participant a total of four times, one
for each block.

### Participants {-}

Participants were undergraduates at the University of Wisconsin-Madison who
completed the experiment in exchange for course credit. We decided to collect
data for 40 first generation participants, and 80 second generation
participants. There were twice as many second generation participants because
each first generation participant was yoked to two second generation
participants, one in each training condition (see Fig. \@ref(fig:gems-team-structures)).

## Results {-}

```{r results}
```

### Training {-}

The purpose of the training procedure was for participants to learn to pay
attention to independent dimensions of the gabor stimuli corresponding
to their training condition. To verify the success of the training procedure, we
measured the extent to which participants selected stimuli in each training
trial corresponding to their training condition.

Participants in both training conditions traversed the same stimulus space
in orthogonal directions (Fig. \@ref(fig:gems-training)A). As expected,
participants selected gems on each trial that were closer to
the most valuable gems on their assigned training landscape. The relationship was
linear, `r r_$training_distance_linear_trial`, with most participants reaching the peak
of the training landscape within 30 trials.

Although participants in both training conditions proceeded in orthogonal
directions, they did not do so equally. Participants trained on bar width were
able to hone in on the optimal stimuli in fewer trials than participants
trained on orientation, `r r_$training_distance_linear_trial_v_instructions`
(Fig. \@ref(fig:gems-training)B).

(ref:gems-training) Results of training procedure. Participants were
randomly assigned to either the orientation or the bar width condition.
**A. Grid positions.** Participants in both training conditions started at
$(0_{ori},0_{bw})$ and traversed the same stimulus space in orthogonal
directions, each progressing toward the optimal stimulus value in the trained
dimension. **B. Distance.** Grid positions were converted to 1D distances
relative to the optimal value in each dimension. **C. Score.** Cumulative
increases in score over trials, indicating that participants were selecting
incrementally more valuable gems over trials.

```{r gems-training, fig.width=10, fig.height=4, fig.cap="(ref:gems-training)"}
gridExtra::grid.arrange(
  training_positions_plot + ggtitle("A. Grid position"),
  training_distance_plot + ggtitle("B. Distance"),
  training_scores_plot + ggtitle("C. Score"),
  nrow = 1
)
```

We also measured the perceptual sensitivity of participants to the different
stimulus dimensions on a trial-by-trial basis. For this analysis, we transformed
the stimulus dimensions of the gems displayed on each trial to be relative to
the optimal value along that dimension. For example, for a participant assigned
to the orientation training condition, gems with orientations at grid points
40 and 60 were both 10 units away from the optimal orientation at 50 units. These
relative values were then centered for each trial such that positive values
indicated that the gem was an improvement in score relative to the
other gems on that trial. Using these transformed values, we fit two models, one
predicting sensitivity to the orientation dimension, and another predicting
sensitivity to the bar width dimension. We included the instruction condition
of the participants in the models. We expected participants in either training
condition to be more sensitive to to their trained dimension than the untrained
dimension.

As expected, participants trained on either stimulus dimension selected
gems on each trial that moved them closer to the optimal stimulus value
in that dimension, orientation: `r r_$training_sensitivity_ori`, bar width:
`r r_$training_sensitivity_sf`. To compare the effectiveness of either training
condition, we transformed the sensitivities to be relative to training
condition. Unlike in the analysis of distance reported above, our
trial-by-trial analysis failed to reveal a difference between training conditions,
`r r_$training_sensitivity_trained_dimension_by_instructions`.

(ref:gems-training-sensitivities) Sensitivity to stimulus dimensions in the training
procedure. Each thin line is the average sensitivity for a single participant.
Thick lines correspond to the predictions of the hierarchical linear model.
**A.** Sensitivity to orientation and bar width separated by training condition.
**B.** Trained dimension collapsed over training condition.

```{r gems-training-sensitivities, fig.width=8, fig.height=3, fig.cap="(ref:gems-training-sensitivities)"}
gridExtra::grid.arrange(
  training_sensitivity_plot + ggtitle("A. Sensitivity to orientation and bar width"),
  training_sensitivity_trained_dimensions_plot +
    ggtitle("B. Trained dimension") +
    # removing y-axis title and legend because of compound plot
    theme(
      axis.title.y = element_blank(),
      legend.position = "none"
    ),
  nrow = 1,
  widths = c(0.66, 0.34)
)
```

The results of the training procedure indicate that participants were
successfully able to discriminate independent dimensions of the gabor patch
stimuli. We now turn to our primary hypothesis of whether participants trained
on independent dimensions were able to coordinate their problem solving
behavior more effectively than participants trained on the same dimension.

### Strategies {-}

To understand whether complementary problem solvers were able to exceed
the problem solving performance of the control strategies, we predicted
performance after 20 trials of the second generation participants to performance
after the same number of trials of the other conditions.

In terms of score after 20 trials, complementary problem solvers were able
to achieve higher scores than both isolated and congruent strategies,
`r r_$strategies_scores_helmert_mod_main`, with no residual difference
between isolated and congruent strategies, `r r_$strategies_scores_helmert_mod_resid`,
(Fig. \@ref(fig:strategies)A).

(ref:strategies) Hill climbing performance. **A.** Scores correspond to heights
on the landscape, and **C.** Distances correspond to Euclidean distances between
current location and peak in the landscape. **B.** and **D.** show these same
measures controlling for starting position.

```{r strategies, fig.width=7, fig.height=7, fig.cap="(ref:strategies)"}
gridExtra::grid.arrange(
  strategies_scores_plot + ggtitle("A. Final score"),
  strategies_relative_scores_plot + ggtitle("B. Relative score"),
  strategies_distance_plot + ggtitle("C. Distance to peak"),
  strategies_relative_distance_plot + ggtitle("D. Relative distance"),
  nrow = 2
)
```

(ref:sensitivities) Sensitivity to trained and untrained dimensions by team strategy.

```{r sensitivities, fig.width=8, fig.cap="(ref:sensitivities)"}
sensitivities_plot
```

## Discussion {-}

In Experiment 2, we sought to test whether teams of participants trained on
different but complementary skills could work together resulting in performance
that was better than independent individuals attempting the same task. We
found support for our hypothesis that problem solving with inheritance can
result in an improvement in performance.

# General Discussion {-}

