---
title: Supplementary materials
output:
  bookdown::pdf_document2:
    toc: true
---
```{r inherited-diversity-config, include=FALSE}
library(knitr)
library(kableExtra)
opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE, results="hide",
               fig.height=4, fig.width=4)
chunks_dir <- "R"
read_chunk(file.path(chunks_dir, "inherited-diversity.R"))
```

# Methods

## Materials

### Gabors

(ref:gabors) Sampled space of possible gabor patches. Gabor patches varied
in orientation and spatial frequency.

```{r gabors, fig.cap="(ref:gabors)"}
draw_image("SimpleHillGems", package = "gems")
```

### Post experiment survey questions

(ref:survey-questions) Questions asked in the post-experiment survey.

```{r survey-questions, results='asis'}
data("Survey")
data_frame(question_text = colnames(Survey)) %>%
  mutate(question_id = c("timestamp", "subj_id", "computer", "age", "gender", paste0("Q", 1:9))) %>%
  filter(grepl("Q", question_id)) %>%
  select(`#` = question_id, `Text` = question_text) %>%
  kable(caption="(ref:survey-questions)") %>%
  column_spec(2, width = "45em")
```

## Procedure

(ref:training) Exact instructions given to participants in either training condition. The instructions differ in the first paragraph and in the sample gabors shown
on the screen.

```{r training, fig.width=8, fig.cap="(ref:training)"}
gridExtra::grid.arrange(
  crotchet::read_image("training_orientation", package = "gems"),
  crotchet::read_image("training_spatial_frequency", package = "gems"),
  nrow = 1
)
```

(ref:training-trial) Training trial procedure. Participants received
feedback in the form of relative changes in their score for each of
the six stimuli on a trial.

```{r training-trial, fig.width=8, fig.cap="(ref:training-trial)"}
gridExtra::grid.arrange(
  crotchet::read_image("training_trial", package = "gems"),
  crotchet::read_image("training_trial_feedback", package = "gems"),
  nrow = 1
)
```

(ref:test-trial) Test trial procedure. Participants received feedback in the
form of the relative change in their score after each selection.

```{r test-trial, fig.width=8, fig.cap="(ref:test-trial)"}
gridExtra::grid.arrange(
  crotchet::read_image("test_trial", package = "gems"),
  crotchet::read_image("test_trial_feedback", package = "gems"),
  nrow = 1
)
```

# Results {-}

## First generation {-}

(ref:gen1-results) Results of first generation problem solvers. Each line is a
participant. **A. Position** on the solution landscape. **B. Distance** between current location and 2D peak. **C. Score** over trials.

```{r gen1-results, fig.width=8, fig.height=10, fig.cap="(ref:gen1-results)"}
gridExtra::grid.arrange(
  gen1_positions_plot + ggtitle("A. Position"),
  gen1_distance_plot + ggtitle("B. Distance"),
  gen1_scores_plot + ggtitle("C. Score"),
  nrow = 3
)
```

(ref:gen1-final-scores) First generation performance by block.

```{r gen1-final-scores, fig.width=6, fig.cap="(ref:gen1-final-scores)"}
gridExtra::grid.arrange(
  gen1_final_scores_plot + ggtitle("A. Final scores"),
  gen1_final_distances_plot + ggtitle("B. Distance to peak"),
  nrow = 1
)
```

## Second generation {-}

(ref:gen2-results) Second generation results.

```{r gen2-results, fig.width=8, fig.height=10, fig.cap="(ref:gen2-results)"}
gridExtra::grid.arrange(
  gen2_positions_plot + ggtitle("A. Position"),
  gen2_distance_plot + ggtitle("B. Distance"),
  gen2_scores_plot + ggtitle("C. Score"),
  nrow = 3
)
```

(ref:gen2-final-scores) Second generation performance by block.

```{r gen2-final-scores, fig.width=6, fig.cap="(ref:gen2-final-scores)"}
gridExtra::grid.arrange(
  gen2_final_scores_plot + ggtitle("A. Final scores"),
  gen2_final_distances_plot + ggtitle("B. Distance to peak"),
  nrow = 1
)
```

## Handoff {-}

(ref:gems-handoff) Handoff between first and second generation participants, showing
isolated, congruent, and complementary conditions.

```{r gems-generation2, fig.width=8, fig.cap="(ref:gems-handoff)"}
Gems %>%
  filter(subj_id == "GEMS120" | ancestor_id == "GEMS120",
         landscape_ix != 0, team_trial <= 40) %>%
  ggplot() +
    aes(team_trial, score, group = subj_id, color = instructions) +
    geom_line(aes(linetype = factor(generation))) +
    facet_wrap("landscape_ix", nrow = 1) +
    t_$theme +
    theme(legend.position = "bottom")
```

## Post experiment survey {-}

```{r survey}
data("Survey")

questions <- data_frame(question_text = colnames(Survey)) %>%
  mutate(question_id = c("timestamp", "subj_id", "computer", "age", "gender", paste0("Q", 1:9))) %>%
  select(question_id, question_text)

get_question_text <- function(question_id) {
  questions[questions$question_id == question_id, ]$question_text
}

colnames(Survey) <- questions$question_id

predictable_plot <- ggplot(Survey) +
  aes(Q7) +
  geom_bar(stat = "count", width = 0.95) +
  scale_x_continuous("", breaks = 1:5, labels = c("1\nRandom", 2, 3, 4, "5\nPredictable")) +
  ggtitle(get_question_text("Q7")) +
  coord_cartesian(expand = FALSE) +
  t_$theme +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
```
